---
title: "R Notebook"
output: html_notebook
---

```{r}
### RESTART ###
library(dplyr)
library(Seurat)
library(patchwork)
library(Matrix)
library(ggplot2)
library(tibble)
library(cowplot)
library(purrr)
wdir <- '~/Jansky/ms_phate/msphate_jansky/msPHATE-Seurat/'

data_name <- 'med_msphate_lvls2345_2k_all.RDS'
med <- readRDS(paste(wdir, data_name, sep = ''))
```


## 1. Comparing 2k and all
```{r}
library(dplyr)
library(Seurat)
library(patchwork)
library(Matrix)
library(ggplot2)
library(tibble)
library(cowplot)
library(purrr)
wdir <- '~/Jansky/ms_phate/msphate_jansky/msPHATE-Seurat/'

data_name <- 'adrenal_medulla_Seurat.RDS.gz'
med <- readRDS(paste(wdir, data_name, sep = ''))

top_varfeats_global <- head(VariableFeatures(med), 10)
LabelPoints(VariableFeaturePlot(med), points = top_varfeats_global, repel = TRUE)
```

Now we'll cluster based only on expression of the most variable genes (initial seurat object unchanged, embedding and cluster identity are metadata belonging to cells, and every gene is stil measured in every cell, so subsetting by genes still keeps all cells)

```{r}
med@commands$RunPCA.RNA@params$features -> jansky_varfeats
VariableFeatures(med) -> med_varfeats
head(med_varfeats %in% jansky_varfeats)
```

Using the same genes as Jansky for our analysis, so now we can proceed

```{r}
GetAssayData(med, 'scale.data') %>% as.data.frame() %>% rownames_to_column(var='genes') %>% filter(genes %in% VariableFeatures(med)) %>% column_to_rownames(var="genes") -> scale_data_2k # Needs pivoting to be filtered by dplyr

write.table(scale_data_2k, file = paste(wdir, 'counts.csv', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(colnames(scale_data_2k), file = paste(wdir, 'cell_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(rownames(scale_data_2k), file = paste(wdir, 'gene_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
# Files written are run through msPHATE in Python
```

Now running on four clustering levels

```{python}
### DO NOT RUN IN R ###
import multiscale_phate as mp
import numpy as np
import pandas as pd
import scprep
import os

wdir = '~/Jansky/ms_phate/msphate_jansky/msPHATE-Seurat'
npca = 20 # Deafault value for RunPCA in Seurat
gran = 0.1
vis_level = 0 # Always zero so points on embedding correspond to individual cells
clus_level = 4

wdir = os.path.expanduser(wdir)
gene_path = os.path.join(wdir, 'gene_names.txt')
cell_path = os.path.join(wdir, 'cell_names.txt')

print('Loading scaled data...')
data_path = os.path.join(wdir, 'counts.csv')
data = scprep.io.load_csv(data_path, cell_axis='column',
                          gene_names=gene_path,
                          cell_names=cell_path)

mp_op = mp.Multiscale_PHATE(n_pca=npca, granularity=gran, random_state=0)
levels = mp_op.fit(data)

print('Generating embedding and assigning clusters...')
embedding, clusters, sizes = mp_op.transform(visualization_level = levels[vis_level],
                                                             cluster_level = levels[-1*clus_level])

print('Exporting embedding and clusters...')
msPHATE_embedding = pd.DataFrame(embedding,
                        index = pd.read_csv(cell_path, header=None).iloc[:, 0].to_list())
msPHATE_embedding.to_csv('msPHATE_embedding.csv', header=False)
msPHATE_clusters = pd.DataFrame(clusters)
msPHATE_clusters.to_csv('msPHATE_clusters.csv', header=False, index=False)

print('\n')
print(data)
print('\n')
print(levels)
print('\n')
print(embedding)
print('\n')
print(set(clusters))
print('Done')
```

```{r}
msPHATE_embedding_global <- read.csv(paste(wdir, 'msPHATE_embedding.csv', sep = ''), header=FALSE, row.names = 1, col.names = c('', 'msPHATE_1', 'msPHATE_2'))
msPHATE_embedding_global <- as.matrix(msPHATE_embedding_global)
med[['msphate_2k']] <- CreateDimReducObject(embeddings = msPHATE_embedding_global, key = 'msPHATE_')
# Only run once - embedding constant for different clustering levels
```

```{r}
msPHATE_clusters_global <- as.matrix(read.csv(paste(wdir, 'msPHATE_clusters.csv', sep = ''), header=FALSE))
msPHATE_clusters_global <- factor(msPHATE_clusters_global)
med$msphate_clusters_lvl5_2k <- msPHATE_clusters_global
# Manually changing through lvls 2/3/4/5 for each msPHATE run
```

Now let's run msPHATE for all data, not just the top 2k most variable genes

```{r}
write.table(GetAssayData(med, 'scale.data'), file = paste(wdir, 'counts.csv', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(colnames(GetAssayData(med, 'scale.data')), file = paste(wdir, 'cell_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(rownames(GetAssayData(med, 'scale.data')), file = paste(wdir, 'gene_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
```

```{r}
msPHATE_embedding_global <- read.csv(paste(wdir, 'msPHATE_embedding.csv', sep = ''), header=FALSE, row.names = 1, col.names = c('', 'msPHATE_1', 'msPHATE_2'))
msPHATE_embedding_global <- as.matrix(msPHATE_embedding_global)
med[['msphate_all']] <- CreateDimReducObject(embeddings = msPHATE_embedding_global, key = 'msPHATE_')
# Only run once - embedding constant for different clustering levels
```

```{r}
msPHATE_clusters_global <- as.matrix(read.csv(paste(wdir, 'msPHATE_clusters.csv', sep = ''), header=FALSE))
msPHATE_clusters_global <- factor(msPHATE_clusters_global)
med$msphate_clusters_lvl5_all <- msPHATE_clusters_global
# Manually changing through lvls 2/3/4/5 for each msPHATE run
```

```{r}
saveRDS(med, file = paste(wdir, 'med_msphate_lvls2345_2k_all.RDS', sep = ''))
```

Let's visualise the clustering and see how using the top 2k most variable genes vs all the genes affects msPHATE clustering compared to UMAP


```{r}
#med$jansky_idents <- Idents(med)


DimPlot(med, reduction = 'msphate', group.by = 'jansky_idents', pt.size = 1) +
DimPlot(med, reduction = 'msphate_all', group.by = 'jansky_idents', pt.size = 1)

DimPlot(med, reduction = 'msphate_all', group.by = 'msphate_clusters_lvl2_all', pt.size = 1) + 
DimPlot(med, reduction = 'msphate_all', group.by = 'msphate_clusters_lvl3_all', pt.size = 1) + 
DimPlot(med, reduction = 'msphate_all', group.by = 'msphate_clusters_lvl4_all', pt.size = 1) + 
DimPlot(med, reduction = 'msphate_all', group.by = 'msphate_clusters_lvl5_all', pt.size = 1)

DimPlot(med, reduction = 'msphate', group.by = 'msphate_clusters_lvl2_2k', pt.size = 1) + 
DimPlot(med, reduction = 'msphate', group.by = 'msphate_clusters_lvl3_2k', pt.size = 1) + 
DimPlot(med, reduction = 'msphate', group.by = 'msphate_clusters_lvl4_2k', pt.size = 1) + 
DimPlot(med, reduction = 'msphate', group.by = 'msphate_clusters_lvl5_2k', pt.size = 1)

umap_check <- med
all_genes <- rownames(GetAssayData(umap_check, 'scale.data'))
umap_check <- ScaleData(umap_check, features = all_genes)
umap_check <- RunPCA(object = umap_check, features = all_genes)
umap_check <- FindNeighbors(object = umap_check)
umap_check <- FindClusters(object = umap_check, resolution = 1.2)
umap_check <- RunUMAP(object = umap_check, features = all_genes)

DimPlot(med, reduction = 'umap', group.by = 'jansky_idents', pt.size = 1) +
DimPlot(med, reduction = 'umap', group.by = 'seurat_clusters', pt.size = 1)
DimPlot(object = umap_check, reduction = "umap", group.by = 'jansky_idents', pt.size = 1) + 
DimPlot(object = umap_check, reduction = "umap", group.by = 'seurat_clusters', pt.size = 1)
#rm(umap_check)
```

## 2. Comparing scaled and unscaled data

Let's check the effect of scaling for the 2k data now. In the intersts of time, only two clustering levels - 3 and 4

```{r}
scale_check <- med
scale_check <- ScaleData(scale_check, do.scale = FALSE)
```

```{r}
GetAssayData(scale_check, 'scale.data') %>% as.data.frame() %>% rownames_to_column(var='genes') %>% filter(genes %in% VariableFeatures(scale_check)) %>% column_to_rownames(var="genes") -> unscaled_data_2k

write.table(unscaled_data_2k, file = paste(wdir, 'counts.csv', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(colnames(unscaled_data_2k), file = paste(wdir, 'cell_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(rownames(unscaled_data_2k), file = paste(wdir, 'gene_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
```

```{r}
msPHATE_embedding_global <- read.csv(paste(wdir, 'msPHATE_embedding.csv', sep = ''), header=FALSE, row.names = 1, col.names = c('', 'msPHATE_1', 'msPHATE_2'))
msPHATE_embedding_global <- as.matrix(msPHATE_embedding_global)
scale_check[['msphate']] <- CreateDimReducObject(embeddings = msPHATE_embedding_global, key = 'msPHATE_')
# Only run once - embedding constant for different clustering levels
```

```{r}
msPHATE_clusters_global <- as.matrix(read.csv(paste(wdir, 'msPHATE_clusters.csv', sep = ''), header=FALSE))
msPHATE_clusters_global <- factor(msPHATE_clusters_global)
scale_check$msphate_clusters_lvl4 <- msPHATE_clusters_global
# Manually changing through lvls 3/4 for each msPHATE run
```

And now for UMAP

```{r}
scale_check <- ScaleData(scale_check, do.scale = FALSE)
scale_check <- RunPCA(object = scale_check)
scale_check <- FindNeighbors(object = scale_check)
scale_check <- FindClusters(object = scale_check, resolution = 1.2)
scale_check <- RunUMAP(object = scale_check, features = VariableFeatures(scale_check))
scale_check <- RunTSNE(object = scale_check, features = VariableFeatures(scale_check))
med <- RunTSNE(object = med, features = VariableFeatures(med))
```

```{r}
DimPlot(scale_check, reduction = 'msphate', group.by = 'msphate_clusters_lvl3', pt.size = 1) +
DimPlot(scale_check, reduction = 'msphate', group.by = 'msphate_clusters_lvl4', pt.size = 1) +
DimPlot(scale_check, reduction = 'msphate', group.by = 'jansky_idents', pt.size = 1)

DimPlot(scale_check, reduction = 'umap', group.by = 'jansky_idents', pt.size = 1) +
DimPlot(med, reduction = 'umap', group.by = 'jansky_idents', pt.size = 1)

DimPlot(scale_check, reduction = 'tsne', group.by = 'jansky_idents', pt.size = 1) +
DimPlot(med, reduction = 'tsne', group.by = 'jansky_idents', pt.size = 1)

DimPlot(scale_check, reduction = 'umap', group.by = 'seurat_clusters', pt.size = 1) +
DimPlot(med, reduction = 'umap', group.by = 'seurat_clusters', pt.size = 1)

#saveRDS(scale_check, file = paste(wdir, 'med_msphate_lvls34_unscaled_2k.RDS'))
```

## 3. Zooming into bridge, connecting chromaffins, and neuroblasts

```{r}
bcn <- subset(med, subset = jansky_idents %in% c('Bridge', 'connecting Chromaffin cells', 'Neuroblasts', 'cycling Neuroblasts', 'late Neuroblasts'))
```

```{r}
bcn <- ScaleData(bcn)
GetAssayData(bcn, 'scale.data') %>% as.data.frame() %>% rownames_to_column(var='genes') %>% filter(genes %in% VariableFeatures(bcn)) %>% column_to_rownames(var="genes") -> scale_data_2k # Needs pivoting to be filtered by dplyr

write.table(scale_data_2k, file = paste(wdir, 'counts.csv', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(colnames(scale_data_2k), file = paste(wdir, 'cell_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(rownames(scale_data_2k), file = paste(wdir, 'gene_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
# Files written are run through msPHATE in Python
```

```{r}
msPHATE_embedding_global <- read.csv(paste(wdir, 'msPHATE_embedding.csv', sep = ''), header=FALSE, row.names = 1, col.names = c('', 'msPHATE_1', 'msPHATE_2'))
msPHATE_embedding_global <- as.matrix(msPHATE_embedding_global)
bcn[['msphate']] <- CreateDimReducObject(embeddings = msPHATE_embedding_global, key = 'msPHATE_')
# Only run once - embedding constant for different clustering levels
```

```{r}
msPHATE_clusters_global <- as.matrix(read.csv(paste(wdir, 'msPHATE_clusters.csv', sep = ''), header=FALSE))
msPHATE_clusters_global <- factor(msPHATE_clusters_global)
bcn$msphate_subclusters_lvl4 <- msPHATE_clusters_global
# Manually changing through lvls 3/4 for each msPHATE run
```

Now we do unscaled

```{r}
bcn_unscaled <- bcn
bcn_unscaled <- ScaleData(bcn_unscaled, do.scale = FALSE)
```

```{r}
GetAssayData(bcn_unscaled, 'scale.data') %>% as.data.frame() %>% rownames_to_column(var='genes') %>% filter(genes %in% VariableFeatures(bcn_unscaled)) %>% column_to_rownames(var="genes") -> unscaled_data_bcn # Needs pivoting to be filtered by dplyr

write.table(unscaled_data_bcn, file = paste(wdir, 'counts.csv', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(colnames(unscaled_data_bcn), file = paste(wdir, 'cell_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(rownames(unscaled_data_bcn), file = paste(wdir, 'gene_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
# Files written are run through msPHATE in Python
```

```{r}
msPHATE_embedding_global <- read.csv(paste(wdir, 'msPHATE_embedding.csv', sep = ''), header=FALSE, row.names = 1, col.names = c('', 'msPHATE_1', 'msPHATE_2'))
msPHATE_embedding_global <- as.matrix(msPHATE_embedding_global)
bcn_unscaled[['msphate']] <- CreateDimReducObject(embeddings = msPHATE_embedding_global, key = 'msPHATE_')
# Only run once - embedding constant for different clustering levels
```

```{r}
msPHATE_clusters_global <- as.matrix(read.csv(paste(wdir, 'msPHATE_clusters.csv', sep = ''), header=FALSE))
msPHATE_clusters_global <- factor(msPHATE_clusters_global)
bcn_unscaled$msphate_subclusters_lvl4 <- msPHATE_clusters_global
# Manually changing through lvls 3/4 for each msPHATE run
```

```{r}
bcn_unscaled <- RunPCA(object = bcn_unscaled)
bcn_unscaled <- FindNeighbors(object = bcn_unscaled)
bcn_unscaled <- FindClusters(object = bcn_unscaled, resolution = 1.2)
bcn_unscaled <- RunUMAP(object = bcn_unscaled, features = VariableFeatures(bcn_unscaled))
bcn_unscaled <- RunTSNE(object = bcn_unscaled, features = VariableFeatures(bcn_unscaled))

bcn <- RunPCA(object = bcn)
bcn <- FindNeighbors(object = bcn)
bcn <- FindClusters(object = bcn, resolution = 1.2)
bcn <- RunUMAP(object = bcn, features = VariableFeatures(bcn))
bcn <- RunTSNE(object = bcn, features = VariableFeatures(bcn))

#saveRDS(bcn_unscaled, file = paste(wdir, ' bcn_msphate_lvls34_unscaled_umap_tsne_2k', sep = ''))
#saveRDS(bcn, file = paste(wdir, 'bcn_msphate_lvls34_unscaled_2k'))
```

```{r}
DimPlot(bcn, reduction = 'msphate', group.by = 'msphate_subclusters_lvl3', pt.size = 1) +
DimPlot(bcn, reduction = 'msphate', group.by = 'msphate_subclusters_lvl4', pt.size = 1) +
DimPlot(bcn, reduction = 'msphate', group.by = 'jansky_idents', pt.size = 1)

DimPlot(bcn_unscaled, reduction = 'msphate', group.by = 'msphate_subclusters_lvl3', pt.size = 1) +
DimPlot(bcn_unscaled, reduction = 'msphate', group.by = 'msphate_subclusters_lvl4', pt.size = 1) +
DimPlot(bcn_unscaled, reduction = 'msphate', group.by = 'jansky_idents', pt.size = 1)

DimPlot(bcn_unscaled, reduction = 'umap', group.by = 'jansky_idents', pt.size = 1) +
DimPlot(bcn, reduction = 'umap', group.by = 'jansky_idents', pt.size = 1)

DimPlot(bcn_unscaled, reduction = 'tsne', group.by = 'jansky_idents', pt.size = 1) +
DimPlot(bcn, reduction = 'tsne', group.by = 'jansky_idents', pt.size = 1)
```
Interesting stuff. Zooming into UMAP-identified clusters with UMAP seems to give bad results

## 3.1. Same, but using corresponding msPHATE clusters

Do I select them manually? Do I find msPHATE clusters overlapping with jansky names of interest? Will the above weirdness remain? I suspect not, as msPHATE clusters are 'designed' to be zoomed into

```{r}
DimPlot(med, reduction = 'msphate', group.by = 'msphate_clusters_lvl2_2k', pt.size = 1) +
DimPlot(med, reduction = 'msphate', group.by = 'msphate_clusters_lvl3_2k', pt.size = 1) +
DimPlot(med, reduction = 'msphate', group.by = 'msphate_clusters_lvl4_2k', pt.size = 1) +
DimPlot(med, reduction = 'msphate', group.by = 'msphate_clusters_lvl5_2k', pt.size = 1) +
DimPlot(med, reduction = 'msphate', group.by = 'jansky_idents', pt.size = 1)
```

```{r}
HoverLocator(DimPlot(med, reduction = 'msphate', group.by = 'msphate_clusters_lvl5_2k', pt.size = 1), information = FetchData(med, vars = 'msphate_clusters_lvl5_2k'))
```

Zooming into combination of 0, 1, 2, 5 (Bridge, connecting + proper Chromaffin cells, Neuroblasts excluding late)
I FORGOT TO RE-FIND VARIABLE FEATURES ABOVE AAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAAA


```{r}
bcn_ms <- subset(med, subset = msphate_clusters_lvl5_2k %in% c('0', '1', '2', '5'))
```

```{r}
bcn_ms <- FindVariableFeatures(bcn_ms)

bcn_ms <- ScaleData(bcn_ms)
GetAssayData(bcn_ms, 'scale.data') %>% as.data.frame() %>% rownames_to_column(var='genes') %>% filter(genes %in% VariableFeatures(bcn_ms)) %>% column_to_rownames(var="genes") -> scaled_data # Needs pivoting to be filtered by dplyr

write.table(scaled_data, file = paste(wdir, 'counts.csv', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(colnames(scaled_data), file = paste(wdir, 'cell_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(rownames(scaled_data), file = paste(wdir, 'gene_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
# Run msPHATE
```

```{r}
msPHATE_embedding_global <- read.csv(paste(wdir, 'msPHATE_embedding.csv', sep = ''), header=FALSE, row.names = 1, col.names = c('', 'msPHATE_1', 'msPHATE_2'))
msPHATE_embedding_global <- as.matrix(msPHATE_embedding_global)
bcn_ms[['msphate']] <- CreateDimReducObject(embeddings = msPHATE_embedding_global, key = 'msPHATE_')
```

```{r}
msPHATE_clusters_global <- as.matrix(read.csv(paste(wdir, 'msPHATE_clusters.csv', sep = ''), header=FALSE))
msPHATE_clusters_global <- factor(msPHATE_clusters_global)
bcn_ms$msphate_subclusters_lvl3 <- msPHATE_clusters_global
# Run msPHATE
```

```{r}
msPHATE_clusters_global <- as.matrix(read.csv(paste(wdir, 'msPHATE_clusters.csv', sep = ''), header=FALSE))
msPHATE_clusters_global <- factor(msPHATE_clusters_global)
bcn_ms$msphate_subclusters_lvl4 <- msPHATE_clusters_global
# Run msPHATE
```

```{r}
bcn_ms_unscaled <- bcn_ms
bcn_ms_unscaled <- FindVariableFeatures(bcn_ms_unscaled)
bcn_ms_unscaled <- ScaleData(bcn_ms_unscaled, do.scale = FALSE)
```

```{r}
GetAssayData(bcn_ms_unscaled, 'scale.data') %>% as.data.frame() %>% rownames_to_column(var='genes') %>% filter(genes %in% VariableFeatures(bcn_ms_unscaled)) %>% column_to_rownames(var="genes") -> unscaled_data # Needs pivoting to be filtered by dplyr

write.table(unscaled_data, file = paste(wdir, 'counts.csv', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(colnames(unscaled_data), file = paste(wdir, 'cell_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(rownames(unscaled_data), file = paste(wdir, 'gene_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
# Run msPHATE
```

```{r}
msPHATE_embedding_global <- read.csv(paste(wdir, 'msPHATE_embedding.csv', sep = ''), header=FALSE, row.names = 1, col.names = c('', 'msPHATE_1', 'msPHATE_2'))
msPHATE_embedding_global <- as.matrix(msPHATE_embedding_global)
bcn_ms_unscaled[['msphate']] <- CreateDimReducObject(embeddings = msPHATE_embedding_global, key = 'msPHATE_')
```

```{r}
msPHATE_clusters_global <- as.matrix(read.csv(paste(wdir, 'msPHATE_clusters.csv', sep = ''), header=FALSE))
msPHATE_clusters_global <- factor(msPHATE_clusters_global)
bcn_ms_unscaled$msphate_subclusters_lvl3 <- msPHATE_clusters_global
# Run msPHATE
```

```{r}
msPHATE_clusters_global <- as.matrix(read.csv(paste(wdir, 'msPHATE_clusters.csv', sep = ''), header=FALSE))
msPHATE_clusters_global <- factor(msPHATE_clusters_global)
bcn_ms_unscaled$msphate_subclusters_lvl4 <- msPHATE_clusters_global
# Run msPHATE
```

```{r}
bcn_ms <- RunPCA(object = bcn_ms)
bcn_ms <- FindNeighbors(object = bcn_ms)
bcn_ms <- FindClusters(object = bcn_ms, resolution = 1.2)
bcn_ms <- RunUMAP(object = bcn_ms, features = VariableFeatures(bcn_ms))
bcn_ms <- RunTSNE(object = bcn_ms, features = VariableFeatures(bcn_ms))

bcn_ms_unscaled <- RunPCA(object = bcn_ms_unscaled)
bcn_ms_unscaled <- FindNeighbors(object = bcn_ms_unscaled)
bcn_ms_unscaled <- FindClusters(object = bcn_ms_unscaled, resolution = 1.2)
bcn_ms_unscaled <- RunUMAP(object = bcn_ms_unscaled, features = VariableFeatures(bcn_ms_unscaled))
bcn_ms_unscaled <- RunTSNE(object = bcn_ms_unscaled, features = VariableFeatures(bcn_ms_unscaled))

#saveRDS(bcn_ms, file = paste(wdir, 'bcn_ms_msphate_lvls34_unscaled_2k'))
#saveRDS(bcn_ms_unscaled, file = paste(wdir, ' bcn_ms_msphate_lvls34_unscaled_umap_tsne_2k', sep = ''))
```

```{r}
DimPlot(bcn_ms, reduction = 'msphate', group.by = 'msphate_subclusters_lvl3', pt.size = 1) +
DimPlot(bcn_ms, reduction = 'msphate', group.by = 'msphate_subclusters_lvl4', pt.size = 1) +
DimPlot(bcn_ms, reduction = 'msphate', group.by = 'jansky_idents', pt.size = 1)

DimPlot(bcn_ms_unscaled, reduction = 'msphate', group.by = 'msphate_subclusters_lvl3', pt.size = 1) +
DimPlot(bcn_ms_unscaled, reduction = 'msphate', group.by = 'msphate_subclusters_lvl4', pt.size = 1) +
DimPlot(bcn_ms_unscaled, reduction = 'msphate', group.by = 'jansky_idents', pt.size = 1)

DimPlot(bcn_ms_unscaled, reduction = 'umap', group.by = 'jansky_idents', pt.size = 1) +
DimPlot(bcn_ms, reduction = 'umap', group.by = 'jansky_idents', pt.size = 1)

DimPlot(bcn_ms_unscaled, reduction = 'tsne', group.by = 'jansky_idents', pt.size = 1) +
DimPlot(bcn_ms, reduction = 'tsne', group.by = 'jansky_idents', pt.size = 1)
```