---
title: "R Notebook"
output: html_notebook
---

```{r}
### RESTART ###
library(dplyr)
library(Seurat)
library(patchwork)
library(Matrix)
library(ggplot2)
library(tibble)
library(cowplot)
library(purrr)
wdir <- '~/Jansky/ms_phate/msphate_jansky/msPHATE-Seurat/'

med <- readRDS(paste(wdir, 'med_msphate_lvls2345_2k_all.RDS', sep = ''))
med_unscaled <- readRDS(paste(wdir, 'med_msphate_lvl4_unscaled_2k.RDS', sep = ''))
bcn <- readRDS(paste(wdir, 'bcn_scaled_lvl4_msphate_umap_tsne_2k.RDS', sep = ''))
bcn_unscaled <- readRDS(paste(wdir, 'bcn_unscaled_lvl4_msphate_umap_tsne_2k.RDS', sep = ''))
bcn_ms <- readRDS(paste(wdir, 'bcn_ms_scaled_lvls34_msphate_umap_tsne_2k.RDS', sep = ''))
bcn_ms_unscaled <- readRDS(paste(wdir, 'bcn_ms_unscaled_lvls34_msphate_umap_tsne_2k.RDS', sep = ''))
```


## 1. Comparing 2k and all
```{r}
library(dplyr)
library(Seurat)
library(patchwork)
library(Matrix)
library(ggplot2)
library(tibble)
library(cowplot)
library(purrr)
wdir <- '~/Jansky/ms_phate/msphate_jansky/msPHATE-Seurat/'

data_name <- 'adrenal_medulla_Seurat.RDS.gz'
med <- readRDS(paste(wdir, data_name, sep = ''))

top_varfeats_global <- head(VariableFeatures(med), 10)
LabelPoints(VariableFeaturePlot(med), points = top_varfeats_global, repel = TRUE)
```

Now we'll cluster based only on expression of the most variable genes (initial seurat object unchanged, embedding and cluster identity are metadata belonging to cells, and every gene is stil measured in every cell, so subsetting by genes still keeps all cells)

```{r}
med@commands$RunPCA.RNA@params$features -> jansky_varfeats
VariableFeatures(med) -> med_varfeats
head(med_varfeats %in% jansky_varfeats)
```

Using the same genes as Jansky for our analysis, so now we can proceed

```{r}
GetAssayData(med, 'scale.data') %>% as.data.frame() %>% rownames_to_column(var='genes') %>% filter(genes %in% VariableFeatures(med)) %>% column_to_rownames(var="genes") -> scale_data_2k # Needs pivoting to be filtered by dplyr

write.table(scale_data_2k, file = paste(wdir, 'counts.csv', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(colnames(scale_data_2k), file = paste(wdir, 'cell_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(rownames(scale_data_2k), file = paste(wdir, 'gene_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
# Files written are run through msPHATE in Python
```

Now running on four clustering levels

```{python}
### DO NOT RUN IN R ###
import multiscale_phate as mp
import numpy as np
import pandas as pd
import scprep
import os

wdir = '~/Jansky/ms_phate/msphate_jansky/msPHATE-Seurat'
npca = 20 # Deafault value for RunPCA in Seurat
gran = 0.1
vis_level = 0 # Always zero so points on embedding correspond to individual cells
clus_level = 4

wdir = os.path.expanduser(wdir)
gene_path = os.path.join(wdir, 'gene_names.txt')
cell_path = os.path.join(wdir, 'cell_names.txt')

print('Loading scaled data...')
data_path = os.path.join(wdir, 'counts.csv')
data = scprep.io.load_csv(data_path, cell_axis='column',
                          gene_names=gene_path,
                          cell_names=cell_path)

mp_op = mp.Multiscale_PHATE(n_pca=npca, granularity=gran, random_state=0)
levels = mp_op.fit(data)

print('Generating embedding and assigning clusters...')
embedding, clusters, sizes = mp_op.transform(visualization_level = levels[vis_level],
                                                             cluster_level = levels[-1*clus_level])

print('Exporting embedding and clusters...')
msPHATE_embedding = pd.DataFrame(embedding,
                        index = pd.read_csv(cell_path, header=None).iloc[:, 0].to_list())
msPHATE_embedding.to_csv('msPHATE_embedding.csv', header=False)
msPHATE_clusters = pd.DataFrame(clusters)
msPHATE_clusters.to_csv('msPHATE_clusters.csv', header=False, index=False)

print('\n')
print(data)
print('\n')
print(levels)
print('\n')
print(embedding)
print('\n')
print(set(clusters))
print('Done')
```

```{r}
msPHATE_embedding_global <- read.csv(paste(wdir, 'msPHATE_embedding.csv', sep = ''), header=FALSE, row.names = 1, col.names = c('', 'msPHATE_1', 'msPHATE_2'))
msPHATE_embedding_global <- as.matrix(msPHATE_embedding_global)
med[['msphate_2k']] <- CreateDimReducObject(embeddings = msPHATE_embedding_global, key = 'msPHATE_')
# Only run once - embedding constant for different clustering levels
```

```{r}
msPHATE_clusters_global <- as.matrix(read.csv(paste(wdir, 'msPHATE_clusters.csv', sep = ''), header=FALSE))
msPHATE_clusters_global <- factor(msPHATE_clusters_global)
med$msphate_clusters_lvl5_2k <- msPHATE_clusters_global
# Manually changing through lvls 2/3/4/5 for each msPHATE run
```

Now let's run msPHATE for all data, not just the top 2k most variable genes

```{r}
write.table(GetAssayData(med, 'scale.data'), file = paste(wdir, 'counts.csv', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(colnames(GetAssayData(med, 'scale.data')), file = paste(wdir, 'cell_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(rownames(GetAssayData(med, 'scale.data')), file = paste(wdir, 'gene_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
```

```{r}
msPHATE_embedding_global <- read.csv(paste(wdir, 'msPHATE_embedding.csv', sep = ''), header=FALSE, row.names = 1, col.names = c('', 'msPHATE_1', 'msPHATE_2'))
msPHATE_embedding_global <- as.matrix(msPHATE_embedding_global)
med[['msphate_all']] <- CreateDimReducObject(embeddings = msPHATE_embedding_global, key = 'msPHATE_')
# Only run once - embedding constant for different clustering levels
```

```{r}
msPHATE_clusters_global <- as.matrix(read.csv(paste(wdir, 'msPHATE_clusters.csv', sep = ''), header=FALSE))
msPHATE_clusters_global <- factor(msPHATE_clusters_global)
med$msphate_clusters_lvl5_all <- msPHATE_clusters_global
# Manually changing through lvls 2/3/4/5 for each msPHATE run
```

```{r}
saveRDS(med, file = paste(wdir, 'med_msphate_lvls2345_2k_all.RDS', sep = ''))
```

Let's visualise the clustering and see how using the top 2k most variable genes vs all the genes affects msPHATE clustering compared to UMAP


```{r}
#med$jansky_idents <- Idents(med)


DimPlot(med, reduction = 'msphate', group.by = 'jansky_idents', pt.size = 1) +
DimPlot(med, reduction = 'msphate_all', group.by = 'jansky_idents', pt.size = 1)

DimPlot(med, reduction = 'msphate_all', group.by = 'msphate_clusters_lvl2_all', pt.size = 1) + 
DimPlot(med, reduction = 'msphate_all', group.by = 'msphate_clusters_lvl3_all', pt.size = 1) + 
DimPlot(med, reduction = 'msphate_all', group.by = 'msphate_clusters_lvl4_all', pt.size = 1) + 
DimPlot(med, reduction = 'msphate_all', group.by = 'msphate_clusters_lvl5_all', pt.size = 1)

DimPlot(med, reduction = 'msphate', group.by = 'msphate_clusters_lvl2_2k', pt.size = 1) + 
DimPlot(med, reduction = 'msphate', group.by = 'msphate_clusters_lvl3_2k', pt.size = 1) + 
DimPlot(med, reduction = 'msphate', group.by = 'msphate_clusters_lvl4_2k', pt.size = 1) + 
DimPlot(med, reduction = 'msphate', group.by = 'msphate_clusters_lvl5_2k', pt.size = 1)

umap_check <- med
all_genes <- rownames(GetAssayData(umap_check, 'scale.data'))
umap_check <- ScaleData(umap_check, features = all_genes)
umap_check <- RunPCA(object = umap_check, features = all_genes)
umap_check <- FindNeighbors(object = umap_check)
umap_check <- FindClusters(object = umap_check, resolution = 1.2)
umap_check <- RunUMAP(object = umap_check, features = all_genes)

DimPlot(med, reduction = 'umap', group.by = 'jansky_idents', pt.size = 1) +
DimPlot(med, reduction = 'umap', group.by = 'seurat_clusters', pt.size = 1)
DimPlot(object = umap_check, reduction = "umap", group.by = 'jansky_idents', pt.size = 1) + 
DimPlot(object = umap_check, reduction = "umap", group.by = 'seurat_clusters', pt.size = 1)
#rm(umap_check)
```

## 2. Comparing scaled and unscaled data

Let's check the effect of scaling for the 2k data now. In the interests of time, only two clustering levels - 3 and 4

```{r}
scale_check <- med
scale_check <- ScaleData(scale_check, do.scale = FALSE)
```

```{r}
GetAssayData(scale_check, 'scale.data') %>% as.data.frame() %>% rownames_to_column(var='genes') %>% filter(genes %in% VariableFeatures(scale_check)) %>% column_to_rownames(var="genes") -> unscaled_data_2k

write.table(unscaled_data_2k, file = paste(wdir, 'counts.csv', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(colnames(unscaled_data_2k), file = paste(wdir, 'cell_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(rownames(unscaled_data_2k), file = paste(wdir, 'gene_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
```

```{r}
msPHATE_embedding_global <- read.csv(paste(wdir, 'msPHATE_embedding.csv', sep = ''), header=FALSE, row.names = 1, col.names = c('', 'msPHATE_1', 'msPHATE_2'))
msPHATE_embedding_global <- as.matrix(msPHATE_embedding_global)
scale_check[['msphate']] <- CreateDimReducObject(embeddings = msPHATE_embedding_global, key = 'msPHATE_')
# Only run once - embedding constant for different clustering levels
```

```{r}
msPHATE_clusters_global <- as.matrix(read.csv(paste(wdir, 'msPHATE_clusters.csv', sep = ''), header=FALSE))
msPHATE_clusters_global <- factor(msPHATE_clusters_global)
scale_check$msphate_clusters_lvl4 <- msPHATE_clusters_global
# Manually changing through lvls 3/4 for each msPHATE run
```

And now for UMAP

```{r}
scale_check <- ScaleData(scale_check, do.scale = FALSE)
scale_check <- RunPCA(object = scale_check)
scale_check <- FindNeighbors(object = scale_check)
scale_check <- FindClusters(object = scale_check, resolution = 1.2)
scale_check <- RunUMAP(object = scale_check, features = VariableFeatures(scale_check))
scale_check <- RunTSNE(object = scale_check, features = VariableFeatures(scale_check))
med <- RunTSNE(object = med, features = VariableFeatures(med))
```

```{r}
DimPlot(scale_check, reduction = 'msphate', group.by = 'msphate_clusters_lvl3', pt.size = 1) +
DimPlot(scale_check, reduction = 'msphate', group.by = 'msphate_clusters_lvl4', pt.size = 1) +
DimPlot(scale_check, reduction = 'msphate', group.by = 'jansky_idents', pt.size = 1)

DimPlot(scale_check, reduction = 'umap', group.by = 'jansky_idents', pt.size = 1) +
DimPlot(med, reduction = 'umap', group.by = 'jansky_idents', pt.size = 1)

DimPlot(scale_check, reduction = 'tsne', group.by = 'jansky_idents', pt.size = 1) +
DimPlot(med, reduction = 'tsne', group.by = 'jansky_idents', pt.size = 1)

DimPlot(scale_check, reduction = 'umap', group.by = 'seurat_clusters', pt.size = 1) +
DimPlot(med, reduction = 'umap', group.by = 'seurat_clusters', pt.size = 1)

#saveRDS(scale_check, file = paste(wdir, 'med_msphate_lvls34_unscaled_2k.RDS'))
```

## 3.1. Zooming into bridge, connecting chromaffins, and neuroblasts using jansky_idents

```{r}
bcn <- subset(med, subset = jansky_idents %in% c('Bridge', 'connecting Chromaffin cells', 'Neuroblasts', 'cycling Neuroblasts', 'late Neuroblasts'))
bcn <- FindVariableFeatures(bcn)
bcn_unscaled <- bcn
bcn <- ScaleData(bcn)
bcn_unscaled <- ScaleData(bcn_unscaled, do.scale = FALSE)
```

```{r}
GetAssayData(bcn, 'scale.data') %>% as.data.frame() %>% rownames_to_column(var='genes') %>% filter(genes %in% VariableFeatures(bcn)) %>% column_to_rownames(var="genes") -> scaled_data # Needs pivoting to be filtered by dplyr

write.table(scaled_data, file = paste(wdir, 'counts.csv', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(colnames(scaled_data), file = paste(wdir, 'cell_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(rownames(scaled_data), file = paste(wdir, 'gene_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
# Run msPHATE
```

```{r}
msPHATE_embedding <- read.csv(paste(wdir, 'msPHATE_embedding.csv', sep = ''), header=FALSE, row.names = 1, col.names = c('', 'msPHATE_1', 'msPHATE_2'))
msPHATE_embedding <- as.matrix(msPHATE_embedding)
bcn[['msphate']] <- CreateDimReducObject(embeddings = msPHATE_embedding, key = 'msPHATE_')
```

```{r}
msPHATE_clusters <- as.matrix(read.csv(paste(wdir, 'msPHATE_clusters.csv', sep = ''), header=FALSE))
msPHATE_clusters_ <- factor(msPHATE_clusters)
bcn$msphate_subclusters_lvl4 <- msPHATE_clusters
```

```{r}
GetAssayData(bcn_unscaled, 'scale.data') %>% as.data.frame() %>% rownames_to_column(var='genes') %>% filter(genes %in% VariableFeatures(bcn_unscaled)) %>% column_to_rownames(var="genes") -> unscaled_data # Needs pivoting to be filtered by dplyr

write.table(unscaled_data, file = paste(wdir, 'counts.csv', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(colnames(unscaled_data), file = paste(wdir, 'cell_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(rownames(unscaled_data), file = paste(wdir, 'gene_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
# Run msPHATE
```

```{r}
msPHATE_embedding <- read.csv(paste(wdir, 'msPHATE_embedding.csv', sep = ''), header=FALSE, row.names = 1, col.names = c('', 'msPHATE_1', 'msPHATE_2'))
msPHATE_embedding <- as.matrix(msPHATE_embedding)
bcn_unscaled[['msphate']] <- CreateDimReducObject(embeddings = msPHATE_embedding, key = 'msPHATE_')
```

```{r}
msPHATE_clusters <- as.matrix(read.csv(paste(wdir, 'msPHATE_clusters.csv', sep = ''), header=FALSE))
msPHATE_clusters <- factor(msPHATE_clusters)
bcn_unscaled$msphate_subclusters_lvl4 <- msPHATE_clusters
```

```{r}
bcn <- RunPCA(object = bcn)
bcn <- FindNeighbors(object = bcn)
bcn <- FindClusters(object = bcn, resolution = 1.2)
bcn <- RunUMAP(object = bcn, features = VariableFeatures(bcn))
bcn <- RunTSNE(object = bcn, features = VariableFeatures(bcn))

bcn_unscaled <- RunPCA(object = bcn_unscaled)
bcn_unscaled <- FindNeighbors(object = bcn_unscaled)
bcn_unscaled <- FindClusters(object = bcn_unscaled, resolution = 1.2)
bcn_unscaled <- RunUMAP(object = bcn_unscaled, features = VariableFeatures(bcn_unscaled))
bcn_unscaled <- RunTSNE(object = bcn_unscaled, features = VariableFeatures(bcn_unscaled))

saveRDS(bcn, file = paste(wdir, 'bcn_scaled_lvl4_msphate_umap_tsne_2k.RDS', sep = ''))
saveRDS(bcn_unscaled, file = paste(wdir, 'bcn_unscaled_lvl4_msphate_umap_tsne_2k.RDS', sep = ''))
```

```{r}
DimPlot(bcn, reduction = 'msphate', group.by = 'msphate_subclusters_lvl4', pt.size = 1) +
  DimPlot(bcn, reduction = 'msphate', group.by = 'jansky_idents', pt.size = 1) +
  FeaturePlot(bcn, reduction = 'msphate', features = 'CC.Difference', cols = c('red', 'green'), pt.size = 1.1, combine=FALSE) + coord_cartesian(xlim = c(-0.02, 0.02), ylim = c(-0.012, 0.016))

DimPlot(bcn_unscaled, reduction = 'msphate', group.by = 'msphate_subclusters_lvl4', pt.size = 1) +
  DimPlot(bcn_unscaled, reduction = 'msphate', group.by = 'jansky_idents', pt.size = 1) +
  FeaturePlot(bcn_unscaled, reduction = 'msphate', features = 'CC.Difference', cols = c('red', 'green'), pt.size = 1.1, combine = FALSE) + coord_cartesian(xlim = c(-0.02, 0.02), ylim = c(-0.012, 0.016))

DimPlot(bcn_unscaled, reduction = 'umap', group.by = 'jansky_idents', pt.size = 1) +
  DimPlot(bcn, reduction = 'umap', group.by = 'jansky_idents', pt.size = 1) + 
  coord_cartesian(xlim = c(-0.02, 0.02), ylim = c(-0.02, 0.025)) +
  FeaturePlot(bcn, reduction = 'umap', features = 'CC.Difference', cols = c('red', 'green'), pt.size = 1.1, combine=FALSE) +
  FeaturePlot(bcn_unscaled, reduction = 'umap', features = 'CC.Difference', cols = c('red', 'green'), pt.size = 1.1, combine=FALSE)

DimPlot(bcn_unscaled, reduction = 'tsne', group.by = 'jansky_idents', pt.size = 1) +
  DimPlot(bcn, reduction = 'tsne', group.by = 'jansky_idents', pt.size = 1) + 
  coord_cartesian(xlim = c(-0.02, 0.02), ylim = c(-0.02, 0.025)) +
  FeaturePlot(bcn, reduction = 'tsne', features = 'CC.Difference', cols = c('red', 'green'), pt.size = 1.1, combine=FALSE) + 
  FeaturePlot(bcn_unscaled, reduction = 'tsne', features = 'CC.Difference', cols = c('red', 'green'), pt.size = 1.1, combine=FALSE)

DimPlot(bcn_unscaled, reduction = 'tsne', group.by = 'jansky_idents', pt.size = 1) +
DimPlot(bcn_unscaled, reduction = 'tsne', group.by = 'seurat_clusters', pt.size = 1) + 
DimPlot(bcn, reduction = 'tsne', group.by = 'jansky_idents', pt.size = 1) +
DimPlot(bcn, reduction = 'tsne', group.by = 'seurat_clusters', pt.size = 1)

DimPlot(bcn_unscaled, reduction = 'umap', group.by = 'jansky_idents', pt.size = 1) +
DimPlot(bcn_unscaled, reduction = 'umap', group.by = 'seurat_clusters', pt.size = 1) +
DimPlot(bcn, reduction = 'umap', group.by = 'jansky_idents', pt.size = 1) +
DimPlot(bcn, reduction = 'umap', group.by = 'seurat_clusters', pt.size = 1)
```


## 3.2. Same, but using corresponding msPHATE clusters

Do I select them manually? Do I find msPHATE clusters overlapping with jansky names of interest? Will the above weirdness remain? I suspect not, as msPHATE clusters are 'designed' to be zoomed into

```{r}
DimPlot(med, reduction = 'msphate', group.by = 'msphate_clusters_lvl2_2k', pt.size = 1) +
DimPlot(med, reduction = 'msphate', group.by = 'msphate_clusters_lvl3_2k', pt.size = 1) +
DimPlot(med, reduction = 'msphate', group.by = 'msphate_clusters_lvl4_2k', pt.size = 1) +
DimPlot(med, reduction = 'msphate', group.by = 'msphate_clusters_lvl5_2k', pt.size = 1) +
DimPlot(med, reduction = 'msphate', group.by = 'jansky_idents', pt.size = 1)
```

```{r}
HoverLocator(DimPlot(med, reduction = 'msphate', group.by = 'msphate_clusters_lvl5_2k', pt.size = 1), information = FetchData(med, vars = 'msphate_clusters_lvl5_2k'))
#DimPlot(med, reduction = 'msphate', group.by = 'msphate_clusters_lvl5_2k', pt.size = 1)
```

Zooming into combination of 0, 1, 2, 5 (Bridge, connecting + proper Chromaffin cells, Neuroblasts)


```{r}
bcn_ms <- subset(med, subset = msphate_clusters_lvl5_2k %in% c('0', '1', '2', '5', '1838', '3920'))
bcn_ms <- subset(bcn_ms, subset = jansky_idents %in% c('Bridge', 'connecting Chromaffin cells', 'Neuroblasts', 'cycling Neuroblasts', 'late Neuroblasts'))
bcn_ms <- FindVariableFeatures(bcn_ms)
bcn_ms_unscaled <- bcn_ms
bcn_ms <- ScaleData(bcn_ms)
bcn_ms_unscaled <- ScaleData(bcn_ms_unscaled, do.scale = FALSE)
```

```{r}
GetAssayData(bcn_ms, 'scale.data') %>% as.data.frame() %>% rownames_to_column(var='genes') %>% filter(genes %in% VariableFeatures(bcn_ms)) %>% column_to_rownames(var="genes") -> scaled_data # Needs pivoting to be filtered by dplyr

write.table(scaled_data, file = paste(wdir, 'counts.csv', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(colnames(scaled_data), file = paste(wdir, 'cell_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(rownames(scaled_data), file = paste(wdir, 'gene_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
# Run msPHATE
```

```{r}
msPHATE_embedding_global <- read.csv(paste(wdir, 'msPHATE_embedding.csv', sep = ''), header=FALSE, row.names = 1, col.names = c('', 'msPHATE_1', 'msPHATE_2'))
msPHATE_embedding_global <- as.matrix(msPHATE_embedding_global)
bcn_ms[['msphate']] <- CreateDimReducObject(embeddings = msPHATE_embedding_global, key = 'msPHATE_')
```

```{r}
msPHATE_clusters_global <- as.matrix(read.csv(paste(wdir, 'msPHATE_clusters.csv', sep = ''), header=FALSE))
msPHATE_clusters_global <- factor(msPHATE_clusters_global)
bcn_ms$msphate_subclusters_lvl3 <- msPHATE_clusters_global
# Did lvls 3 and 4 manually
```

```{r}
GetAssayData(bcn_ms_unscaled, 'scale.data') %>% as.data.frame() %>% rownames_to_column(var='genes') %>% filter(genes %in% VariableFeatures(bcn_ms_unscaled)) %>% column_to_rownames(var="genes") -> unscaled_data # Needs pivoting to be filtered by dplyr

write.table(unscaled_data, file = paste(wdir, 'counts.csv', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(colnames(unscaled_data), file = paste(wdir, 'cell_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(rownames(unscaled_data), file = paste(wdir, 'gene_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
# Run msPHATE
```

```{r}
msPHATE_embedding_global <- read.csv(paste(wdir, 'msPHATE_embedding.csv', sep = ''), header=FALSE, row.names = 1, col.names = c('', 'msPHATE_1', 'msPHATE_2'))
msPHATE_embedding_global <- as.matrix(msPHATE_embedding_global)
bcn_ms_unscaled[['msphate']] <- CreateDimReducObject(embeddings = msPHATE_embedding_global, key = 'msPHATE_')
```

```{r}
msPHATE_clusters_global <- as.matrix(read.csv(paste(wdir, 'msPHATE_clusters.csv', sep = ''), header=FALSE))
msPHATE_clusters_global <- factor(msPHATE_clusters_global)
bcn_ms_unscaled$msphate_subclusters_lvl3 <- msPHATE_clusters_global
# Did lvls 3 and 4 manually
```

```{r}
bcn_ms <- RunPCA(object = bcn_ms)
bcn_ms <- FindNeighbors(object = bcn_ms)
bcn_ms <- FindClusters(object = bcn_ms, resolution = 1.2)
bcn_ms <- RunUMAP(object = bcn_ms, features = VariableFeatures(bcn_ms))
bcn_ms <- RunTSNE(object = bcn_ms, features = VariableFeatures(bcn_ms))

bcn_ms_unscaled <- RunPCA(object = bcn_ms_unscaled)
bcn_ms_unscaled <- FindNeighbors(object = bcn_ms_unscaled)
bcn_ms_unscaled <- FindClusters(object = bcn_ms_unscaled, resolution = 1.2)
bcn_ms_unscaled <- RunUMAP(object = bcn_ms_unscaled, features = VariableFeatures(bcn_ms_unscaled))
bcn_ms_unscaled <- RunTSNE(object = bcn_ms_unscaled, features = VariableFeatures(bcn_ms_unscaled))

saveRDS(bcn_ms, file = paste(wdir, 'bcn_ms_scaled_lvls34_msphate_umap_tsne_2k.RDS', sep = ''))
saveRDS(bcn_ms_unscaled, file = paste(wdir, 'bcn_ms_unscaled_lvls34_msphate_umap_tsne_2k.RDS', sep = ''))
```

```{r}
#DimPlot(bcn_ms, reduction = 'msphate', group.by = 'msphate_subclusters_lvl3', pt.size = 1) +
DimPlot(bcn_ms, reduction = 'msphate', group.by = 'msphate_subclusters_lvl4', pt.size = 1) +
DimPlot(bcn_ms, reduction = 'msphate', group.by = 'jansky_idents', pt.size = 1) +
FeaturePlot(bcn_ms, reduction = 'msphate', features = 'CC.Difference', cols = c('red', 'green'), pt.size = 1.1, combine=FALSE) + coord_cartesian(xlim = c(-0.02, 0.02), ylim = c(-0.015, 0.02))

#DimPlot(bcn_ms_unscaled, reduction = 'msphate', group.by = 'msphate_subclusters_lvl3', pt.size = 1) +
DimPlot(bcn_ms_unscaled, reduction = 'msphate', group.by = 'msphate_subclusters_lvl4', pt.size = 1) +
DimPlot(bcn_ms_unscaled, reduction = 'msphate', group.by = 'jansky_idents', pt.size = 1) +
FeaturePlot(bcn_ms_unscaled, reduction = 'msphate', features = 'CC.Difference', cols = c('red', 'green'), pt.size = 1.1, combine = FALSE) + coord_cartesian(xlim = c(-0.015, 0.02), ylim = c(-0.015, 0.015))

DimPlot(bcn_ms_unscaled, reduction = 'umap', group.by = 'jansky_idents', pt.size = 1) +
DimPlot(bcn_ms, reduction = 'umap', group.by = 'jansky_idents', pt.size = 1) + 
coord_cartesian(xlim = c(-0.02, 0.02), ylim = c(-0.02, 0.025)) +
FeaturePlot(bcn_ms, reduction = 'umap', features = 'CC.Difference', cols = c('red', 'green'), pt.size = 1.1, combine=FALSE) +
FeaturePlot(bcn_ms_unscaled, reduction = 'umap', features = 'CC.Difference', cols = c('red', 'green'), pt.size = 1.1, combine=FALSE)

DimPlot(bcn_ms_unscaled, reduction = 'tsne', group.by = 'jansky_idents', pt.size = 1) +
DimPlot(bcn_ms, reduction = 'tsne', group.by = 'jansky_idents', pt.size = 1) + 
coord_cartesian(xlim = c(-0.02, 0.02), ylim = c(-0.02, 0.025)) +
FeaturePlot(bcn_ms, reduction = 'tsne', features = 'CC.Difference', cols = c('red', 'green'), pt.size = 1.1, combine=FALSE) + 
FeaturePlot(bcn_ms_unscaled, reduction = 'tsne', features = 'CC.Difference', cols = c('red', 'green'), pt.size = 1.1, combine=FALSE)

DimPlot(bcn_ms_unscaled, reduction = 'tsne', group.by = 'jansky_idents', pt.size = 1) +
DimPlot(bcn_ms_unscaled, reduction = 'tsne', group.by = 'seurat_clusters', pt.size = 1) + 
DimPlot(bcn_ms, reduction = 'tsne', group.by = 'jansky_idents', pt.size = 1) +
DimPlot(bcn_ms, reduction = 'tsne', group.by = 'seurat_clusters', pt.size = 1)

DimPlot(bcn_ms_unscaled, reduction = 'umap', group.by = 'jansky_idents', pt.size = 1) +
DimPlot(bcn_ms_unscaled, reduction = 'umap', group.by = 'seurat_clusters', pt.size = 1) +
DimPlot(bcn_ms, reduction = 'umap', group.by = 'jansky_idents', pt.size = 1) +
DimPlot(bcn_ms, reduction = 'umap', group.by = 'seurat_clusters', pt.size = 1)
```

## 4. Analysing plots

### Zooming into jansky_idents

#### msPHATE Scaled

```{r}
DimPlot(bcn, reduction = 'msphate', group.by = 'msphate_subclusters_lvl4', pt.size = 1) +
  DimPlot(bcn, reduction = 'msphate', group.by = 'jansky_idents', pt.size = 1) +
  FeaturePlot(bcn, reduction = 'msphate', features = 'CC.Difference', cols = c('red', 'green'), pt.size = 1.1, combine=FALSE) + coord_cartesian(xlim = c(-0.02, 0.02), ylim = c(-0.012, 0.016))

DimPlot(bcn, reduction = 'msphate', group.by = 'jansky_idents', split.by = 'timepoint2', pt.size = 1)
```
- Late N separated from B+cC+N
- Cycling N separated from late N - threshold selection by Jansky?
- Unusual cells appearing at 11pcw, also gap in late Neuroblasts (experimental thing?)

#### msPHATE Unscaled

```{r}
DimPlot(bcn_unscaled, reduction = 'msphate', group.by = 'msphate_subclusters_lvl4', pt.size = 1) +
  DimPlot(bcn_unscaled, reduction = 'msphate', group.by = 'jansky_idents', pt.size = 1) +
  FeaturePlot(bcn_unscaled, reduction = 'msphate', features = 'CC.Difference', cols = c('red', 'green'), pt.size = 1.1, combine = FALSE) + coord_cartesian(xlim = c(-0.02, 0.02), ylim = c(-0.012, 0.016))

DimPlot(bcn_unscaled, reduction = 'msphate', group.by = 'jansky_idents', split.by = 'timepoint2', pt.size = 1)
```

- Cell cycle appears smaller in magnitude (cycling Neuroblasts better correpond to N and late N than in scaled data fed into msPHATE)
- Same unusual patch at 11pcw (confirmed by jansky_idents group.by), just more spread when using unscaled data
- Unscaled generally more spread and more rep of magnitude


#### tSNE unscaled/scaled

```{r}
DimPlot(bcn_unscaled, reduction = 'tsne', group.by = 'jansky_idents', pt.size = 1) +
  DimPlot(bcn, reduction = 'tsne', group.by = 'jansky_idents', pt.size = 1) + 
  coord_cartesian(xlim = c(-0.02, 0.02), ylim = c(-0.02, 0.025)) +
  FeaturePlot(bcn_unscaled, reduction = 'tsne', features = 'CC.Difference', cols = c('red', 'green'), pt.size = 1.1, combine=FALSE) +
  FeaturePlot(bcn, reduction = 'tsne', features = 'CC.Difference', cols = c('red', 'green'), pt.size = 1.1, combine=FALSE)
```

- Cycling cells are more spread throiugh data in unscaled, more separate in scaled

```{r}
DimPlot(bcn_unscaled, reduction = 'tsne', group.by = 'jansky_idents', pt.size = 1) + 
  DimPlot(bcn_unscaled, reduction = 'tsne', group.by = 'msphate_subclusters_lvl4', pt.size = 1) +
  DimPlot(bcn_unscaled, reduction = 'tsne', group.by = 'seurat_clusters', pt.size = 1)

DimPlot(bcn_unscaled, reduction = 'tsne', group.by = 'msphate_subclusters_lvl4', split.by = 'timepoint2', pt.size = 1)
```

- msPHATE clusters correspond well to tSNE embedding (not clustering)
- Time course visible, but same info better on msPHATE plot

#### UMAP unscaled/scaled

```{r}
DimPlot(bcn_unscaled, reduction = 'umap', group.by = 'jansky_idents', pt.size = 1) +
  DimPlot(bcn, reduction = 'umap', group.by = 'jansky_idents', pt.size = 1) + 
  coord_cartesian(xlim = c(-0.02, 0.02), ylim = c(-0.02, 0.025)) +
  FeaturePlot(bcn_unscaled, reduction = 'umap', features = 'CC.Difference', cols = c('red', 'green'), pt.size = 1.1, combine=FALSE) +
  FeaturePlot(bcn, reduction = 'umap', features = 'CC.Difference', cols = c('red', 'green'), pt.size = 1.1, combine=FALSE)
```

- No difference between scaled/unscaled
- Cycling cells appear like scaled data in other methods (tightly separated group)

```{r}
DimPlot(bcn_unscaled, reduction = 'umap', group.by = 'jansky_idents', split.by = 'timepoint2', pt.size = 1)
```
- Very good for time course

### Zooming into msPHATE clusters corresponding to jansky_idents

#### msPHATE

```{r}
#DimPlot(bcn_ms, reduction = 'msphate', group.by = 'msphate_subclusters_lvl3', pt.size = 1) +
DimPlot(bcn_ms, reduction = 'msphate', group.by = 'msphate_subclusters_lvl4', pt.size = 1) +
DimPlot(bcn_ms, reduction = 'msphate', group.by = 'jansky_idents', pt.size = 1) +
FeaturePlot(bcn_ms, reduction = 'msphate', features = 'CC.Difference', cols = c('red', 'green'), pt.size = 1.1, combine=FALSE) + coord_cartesian(xlim = c(-0.02, 0.02), ylim = c(-0.015, 0.02))

DimPlot(bcn_ms_unscaled, reduction = 'msphate', group.by = 'msphate_subclusters_lvl3', pt.size = 1) +
#DimPlot(bcn_ms_unscaled, reduction = 'msphate', group.by = 'msphate_subclusters_lvl4', pt.size = 1) +
DimPlot(bcn_ms_unscaled, reduction = 'msphate', group.by = 'jansky_idents', pt.size = 1) +
FeaturePlot(bcn_ms_unscaled, reduction = 'msphate', features = 'CC.Difference', cols = c('red', 'green'), pt.size = 1.1, combine = FALSE) + coord_cartesian(xlim = c(-0.015, 0.02), ylim = c(-0.015, 0.015))
```

- Scaled separates cycling cells, unscaled does not

```{r}
#DimPlot(bcn_ms_unscaled, reduction = 'msphate', group.by = 'msphate_subclusters_lvl3', pt.size = 1) +
DimPlot(bcn_ms_unscaled, reduction = 'msphate', group.by = 'msphate_subclusters_lvl4', pt.size = 1) +
DimPlot(bcn_ms_unscaled, reduction = 'msphate', group.by = 'jansky_idents', pt.size = 1) +
FeaturePlot(bcn_ms_unscaled, reduction = 'msphate', features = 'CC.Difference', cols = c('red', 'green'), pt.size = 1.1, combine = FALSE) + coord_cartesian(xlim = c(-0.015, 0.02), ylim = c(-0.015, 0.015))
```

- Here, subcluster 1 looks good to look into as it's a connector of many other types of cells

```{r}
DimPlot(bcn_ms, reduction = 'msphate', group.by = 'timepoint2', pt.size = 1)
```
- Good plot for time course
- Will use lvl4 for looking into subclusters, instead of zooming into subclusters

#### tSNE

```{r}
DimPlot(bcn_ms_unscaled, reduction = 'tsne', group.by = 'jansky_idents', pt.size = 1) +
DimPlot(bcn_ms, reduction = 'tsne', group.by = 'jansky_idents', pt.size = 1) + 
coord_cartesian(xlim = c(-0.02, 0.02), ylim = c(-0.02, 0.025)) +
FeaturePlot(bcn_ms_unscaled, reduction = 'tsne', features = 'CC.Difference', cols = c('red', 'green'), pt.size = 1.1, combine=FALSE) +
FeaturePlot(bcn_ms, reduction = 'tsne', features = 'CC.Difference', cols = c('red', 'green'), pt.size = 1.1, combine=FALSE)
```

- Scaling still pulls out cycling cells
- Still a poor plot to split by timepoint

#### UMAP

```{r}
DimPlot(bcn_ms_unscaled, reduction = 'umap', group.by = 'jansky_idents', pt.size = 1) +
DimPlot(bcn_ms, reduction = 'umap', group.by = 'jansky_idents', pt.size = 1) + 
coord_cartesian(xlim = c(-0.02, 0.02), ylim = c(-0.02, 0.025)) +
FeaturePlot(bcn_ms_unscaled, reduction = 'umap', features = 'CC.Difference', cols = c('red', 'green'), pt.size = 1.1, combine=FALSE) +
FeaturePlot(bcn_ms, reduction = 'umap', features = 'CC.Difference', cols = c('red', 'green'), pt.size = 1.1, combine=FALSE)
```

- Still no difference (like scaled always), still good for time course, cycling N separated into two- Why?

```{r}
DimPlot(bcn_ms_unscaled, reduction = 'umap', group.by = 'timepoint2', pt.size = 1)
DimPlot(bcn_ms_unscaled, reduction = 'umap', group.by = 'msphate_subclusters_lvl4', pt.size = 1)
```

- Subclusters 1 and 10: 10 was budding off to chromaffins on an earlier plot (WAS IT??), 1 is still interesting because it's on the neuroblast trajectory but something makes it separate (PC plot might show? Marker genes?)
- Same conclusions for cycling, time course
