---
title: "Integrating msPHATE, UMAP, and tSNE to explore scRNA-seq data"
output: pdf_document
date: "2022-12-02"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The deal

Here is the deal I am proposing. A good workflow is to do global clustering on published, normalised, scaled data with msPHATE, manually merge clusters if necessary based on heatmaps, subset a global msPHATE cluster of interest into a separate Seurat object, recluster that with Louvain algorithm in Seurat, and visualise the result with tSNE.

Considerations: - Split initial msPHATE clusters? Can define gene modules to do this automatically... - Need to zoom into msPHATE global clusters because zooming into one UMAP cluster was bad once. See if umap/tsne clusters are always unfit for zooming - Does scaling need to be rerun when reclustering? Check literature (they 're-embedded' the)

There is also another potential arm to this project. I want to functionally annotate the clusters I find, either with the above msPHATE-UMAP-tSNE workflow or jus the clusters from a study.

## The process

### Running msPHATE
```{r libraries}
library(dplyr)
library(Seurat)
library(patchwork)
library(Matrix)
library(ggplot2)
```

```{r global_prework}
wdir <- '~/Jansky/ms_phate/msphate_jansky/msPHATE-Seurat/'
data_name <- 'adrenal_medulla_Seurat.RDS.gz'
med <- readRDS(paste(wdir, data_name, sep = ''))

GetAssayData(med, 'scale.data') %>% as.data.frame() %>% rownames_to_column(var='genes') %>% filter(genes %in% VariableFeatures(med)) %>% column_to_rownames(var="genes") -> varfeature_scaleddata
write.table(varfeature_scaleddata, file = paste(wdir, 'counts.csv', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(colnames(varfeature_scaleddata), file = paste(wdir, 'cell_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(rownames(varfeature_scaleddata), file = paste(wdir, 'gene_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)



write.table(GetAssayData(med, 'scale.data'), file = paste(wdir, 'counts.csv', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(colnames(GetAssayData(med, 'scale.data')), file = paste(wdir, 'cell_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(rownames(GetAssayData(med, 'scale.data')), file = paste(wdir, 'gene_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
```

Insert explanation of above

```{python msphate}
### DO NOT RUN IN Rmd!!! Python chunks a bit funny in R
# Script is bloated, will fix later

import multiscale_phate as mp
import numpy as np
import pandas as pd
import scprep
import os
import matplotlib.pyplot as plt
import pickle

# Seurat: unsure how to find out QC from seurat object
# Assuming tutorial defaults
# Seurat: FindNeighbours used 20 PC's
# Seurat: FindClusters (resolution = 1.2)
# Shunya's report: min 500 genes, min 1000 counts, max 2.5% mt - where are QC params from?
wdir = '~/Jansky/ms_phate/msphate_jansky/msPHATE-Seurat'
do_scaling = False
min_reads = 200
mincells = 3
npca = 20
gran = 0.1
run_multiple_embeddings=False
find_expression = False
multiple_spread = 1
vis_level = 0
clus_level = 3
zoom_cluster = 1
marker_dict = {'SCPs': ['SOX10', 'PLP1', 'ERBB3', 'MPZ', 'FOXD3'],
               'Neuroblasts': ['ISL1', 'STMN2', 'NEFM'],
               'Chromaffin cells': ['DBH', 'PHOX2B']}

names = list(marker_dict.keys())
genes = []
for i in marker_dict:
    for j in marker_dict[i]:
        genes.append(j)
do_pickling = False
generate_plots = False
do_zoom=False
if run_multiple_embeddings:
    generate_plots = True

wdir = os.path.expanduser(wdir)
gene_path = os.path.join(wdir, 'gene_names.txt')
cell_path = os.path.join(wdir, 'cell_names.txt')

if do_scaling:
    print('Loading counts matrix...')
    data_path = os.path.join(wdir, 'counts.mtx')
    data = scprep.io.load_mtx(data_path, cell_axis='column',
                              gene_names=gene_path,
                              cell_names=cell_path,
                              sparse=True)

    print('Filtering and normalising...')
    data = scprep.filter.filter_library_size(data, cutoff=min_reads,
                                             keep_cells='above')
    data = scprep.filter.filter_rare_genes(data, min_cells = mincells)
    data = scprep.normalize.library_size_normalize(data)
    data = np.sqrt(data)
    # Can QC'd and normalised counts matrix be imported from Seurat?
else:
    print('Loading scaled data...')
    data_path = os.path.join(wdir, 'counts.csv')
    data = scprep.io.load_csv(data_path, cell_axis='column',
                              gene_names=gene_path,
                              cell_names=cell_path)

mp_op = mp.Multiscale_PHATE(n_pca=npca, granularity=gran, random_state=0)
levels = mp_op.fit(data)
if generate_plots:
    ax = plt.plot(mp_op.gradient)
    ax = plt.scatter(levels, mp_op.gradient[levels], c = 'r', s=100)
    f_dir = os.path.expanduser(wdir)
    f_name = os.path.join(f_dir, 'levels.png')
    plt.savefig(fname=f_name)
    plt.close()

if generate_plots:
    tree = mp_op.build_tree()
    tree_clusters = mp_op.get_tree_clusters(levels[-1*clus_level])
    scprep.plot.scatter3d(tree, c = tree_clusters, s= 50,
                          fontsize=16, ticks=False, figsize=(10,10))
    f_dir = os.path.expanduser(wdir)
    f_name = os.path.join(f_dir, 'tree.png')
    plt.savefig(fname=f_name)
    plt.close()

print('Generating embedding(s)...')
if run_multiple_embeddings:
    for i in range(vis_level, round(len(levels)/2)+multiple_spread, 2):
        for j in [-1*clus_level, -1*clus_level+multiple_spread, -1*clus_level-multiple_spread]:
            embedding, clusters, sizes = mp_op.transform(visualization_level = levels[i],
                                                         cluster_level = levels[j])
            if generate_plots:
                scprep.plot.scatter2d(embedding, s = 100*np.sqrt(sizes), c = clusters,
                                      fontsize=16, ticks=False,
                                      label_prefix="Multiscale PHATE", figsize=(10,8))
                f_dir = os.path.expanduser(wdir)
                custom_name = 'embedding_vis' + str(i) + '_clus' + str(j)+ '.png'
                f_name = os.path.join(f_dir, custom_name)
                plt.savefig(fname=f_name)
                plt.close()
else:
    if do_zoom:
        embedding, clusters, sizes = mp_op.transform(visualization_level = levels[vis_level],
                                                     cluster_level = levels[-1*clus_level],
                                                     coarse_cluster_level = levels[-1*clus_level],
                                                     coarse_cluster=zoom_cluster)
    else:
        embedding, clusters, sizes = mp_op.transform(visualization_level = levels[vis_level],
                                                             cluster_level = levels[-1*clus_level])
        if generate_plots:
            scprep.plot.scatter2d(embedding, s = 100*np.sqrt(sizes), c = clusters,
                                          fontsize=16, ticks=False,
                                          label_prefix="Multiscale PHATE", figsize=(10,8))
            f_dir = os.path.expanduser(wdir)
            custom_name = 'embedding_vis' + str(vis_level) + '_clus' + str(clus_level)+ '.png'
            f_name = os.path.join(f_dir, custom_name)
            plt.savefig(fname=f_name)
            plt.close()

if find_expression:
    print('Finding expression')
    expression = pd.DataFrame()
    for i in range(len(genes)):
        expression[genes[i]] = mp_op.get_expression(data[genes[i]].values,
        visualization_level = levels[vis_level])

if do_pickling:
    print('Pickling important files...')
    keep = [embedding, clusters, sizes, data]
    keep_names = ['embedding', 'clusters', 'sizes', 'data', 'expression']
    for i in range(len(keep)):
        pickle_out = open(keep_names[i] + '.pickle', 'wb')
        pickle.dump(keep[i], pickle_out)
        pickle_out.close()

print('Exporting embedding and clusters...')
msPHATE_embedding = pd.DataFrame(embedding,
                        index = pd.read_csv(cell_path, header=None).iloc[:, 0].to_list())
msPHATE_embedding.to_csv('msPHATE_embedding.csv', header=False)
msPHATE_clusters = pd.DataFrame(clusters)
msPHATE_clusters.to_csv('msPHATE_clusters.csv', header=False, index=False)

print(data)
print(levels)
print('Done')
```

Insert explanation of above

-   vis_level = 0 (always!)
-   clus_level = 3 (7 clusters, 2 gives 3 clusters)

```{r global_postwork}
msPHATE_embedding <- read.csv(paste(wdir, 'msPHATE_embedding.csv', sep = ''), header=FALSE, row.names = 1, col.names = c('', 'msPHATE_1', 'msPHATE_2'))
msPHATE_embedding <- as.matrix(msPHATE_embedding)
med[['msphate']] <- CreateDimReducObject(embeddings = msPHATE_embedding, key = 'msPHATE_')

msPHATE_clusters <- as.matrix(read.csv(paste(wdir, 'msPHATE_clusters.csv', sep = ''), header=FALSE))
msPHATE_clusters <- factor(msPHATE_clusters)
med$msphate_clusters <- msPHATE_clusters

med$jansky_idents <- Idents(med)
Idents(med) <- 'msphate_clusters'
DimPlot(med, reduction = 'msphate', pt.size = 1)
#saveRDS(med, file = paste(wdir, 'med_msphate_global.RDS', sep = '')) # Time-consuming step
```
#### Dealing with its consequences
```{r global_cluster_renaming}
med$msphate_orig_clusters <- Idents(med)
msphate_new_idents <- c('1', '2', '3', '4', '5', '6')
names(msphate_new_idents) <- levels(med)
med <- RenameIdents(med, msphate_new_idents)
med$msphate_clusters_premerge <- Idents(med)
DimPlot(med, reduction = 'msphate', pt.size = 1)
```

```{r global_cluster_renaming}
med <- FindVariableFeatures(med)
top_de <- head(VariableFeatures(med), 10)
LabelPoints(VariableFeaturePlot(med), points = top_de, repel = TRUE)

msphate_varmarkers_global1 <- FindAllMarkers(med, features = VariableFeatures(med))
msphate_varmarkers_global1 %>%
  group_by(cluster) %>%
  top_n(n = 7, wt = avg_log2FC) -> top_de
DoHeatmap(med, features = top_de$gene) + NoLegend()
DimPlot(med, reduction = 'msphate', pt.size = 1)
```
- merging 4 and 5

```{r global_cluster_merging}
msphate_new_idents <- c('1', '2', '1', '3', '3', '4')
names(msphate_new_idents) <- levels(med)
med <- RenameIdents(med, msphate_new_idents)
med$msphate_clusters <- Idents(med)
DimPlot(med, reduction = 'msphate', pt.size = 1)

Idents(med) <- 'msphate_clusters' # Necessary to run findallmarkers correctly
msphate_varmarkers_global2 <- FindAllMarkers(med, features = VariableFeatures(med))
msphate_varmarkers_global2 %>%
  group_by(cluster) %>%
  top_n(n = 7, wt = avg_log2FC) -> top_de
DoHeatmap(med, features = top_de$gene) + NoLegend()
```

#### Comparing to Jansky

- We can now do a few interesting things
- How well do the clusters correlate to jansky ones?

```{r global_msphate_play}
DimPlot(med, reduction = 'msphate', group.by = 'jansky_idents', pt.size = 1.5)

FeaturePlot(med, reduction = 'msphate', features = 'CC.Difference', cols = c('red', 'green'), pt.size = 1.5) + 
  coord_cartesian(xlim = c(-0.01, 0.02), ylim = c(-0.01, 0.02)) +
FeaturePlot(med, reduction = 'umap', features = 'CC.Difference', cols = c('red', 'green'), pt.size = 1.5)
```
- 1 - Bridge and some late SCPs, 2 - Neuroblasts and connecting + most late Chromaffins, 3 - most SCPs, 4 - Chromaffins, 5 - some late Chromaffins (fig 5 to fig 7)
- Cycling cells cluster much better on a umap
- PRBLEM


### Reclustering

- We can now zoom into a cluster of interest
- This is specifically a feature of msPHATE and the reason for using it at all
- Only 2 is uncertain
- We'll zoom into ("re-embed") global msPHATE cluster 2

#### Using msPHATE

- Could recluster using msPHATE

```{r ms_clus2_1}
clus <- subset(med, subset = msphate_clusters %in% c('2')) # Syntax allows multiple clusters
scaled_data_clus2 <- GetAssayData(clus, 'scale.data')
wdir = '~/Jansky/ms_phate/msphate_jansky/msPHATE-Seurat/'
write.table(scaled_data_clus2, 
            file = paste(wdir, 'counts.csv', sep = ''), 
            sep = ',', row.names = FALSE, col.names = FALSE)
write.table(rownames(scaled_data_clus2), file = paste(wdir, 'gene_names.txt', sep = ''), 
            sep = ',', row.names = FALSE, col.names = FALSE)
write.table(colnames(scaled_data_clus2), file = paste(wdir, 'cell_names.txt', sep = ''), 
            sep = ',', row.names = FALSE, col.names = FALSE)
```

```{python ms_clus2_2}
## Python section
```

```{r ms_clus2_3}
msPHATE_embedding <- read.csv(paste(wdir, 'msPHATE_embedding.csv', sep = ''), 
                              header=FALSE,
                              row.names = 1, 
                              col.names = c('', 'msPHATE_1', 'msPHATE_2'))
msPHATE_embedding <- as.matrix(msPHATE_embedding)
clus[['msphate']] <- CreateDimReducObject(embeddings = msPHATE_embedding, key = 'msPHATE_')
msPHATE_clusters <- as.matrix(read.csv(paste(wdir, 'msPHATE_clusters.csv', sep = ''), header=FALSE))
msPHATE_clusters <- factor(msPHATE_clusters)
clus$msphate_clusters <- msPHATE_clusters
Idents(clus) <- 'msphate_clusters'

DimPlot(clus, reduction = 'msphate', pt.size = 1)
```

```{r ms_clus2_4}
clus$msphate_orig_clusters <- Idents(clus)
msphate_new_idents <- c('1', '2', '3', '4', '5', '6', '7')
names(msphate_new_idents) <- levels(clus)
clus <- RenameIdents(clus, msphate_new_idents)
clus$msphate_clusters <- Idents(clus)
Idents(clus) <- 'msphate_clusters'
DimPlot(clus, reduction = 'msphate', pt.size = 1)

clus <- FindVariableFeatures(clus)
top_de <- head(VariableFeatures(clus), 10)
LabelPoints(VariableFeaturePlot(clus), points = top_de, repel = TRUE)

msphate_varmarkers_clus2_1 <- FindAllMarkers(clus, features = VariableFeatures(clus))
msphate_varmarkers_clus2_1 %>%
  group_by(cluster) %>%
  top_n(n = 7, wt = avg_log2FC) -> top_de
DoHeatmap(clus, features = top_de$gene) + NoLegend()
```
- Merge 4 and 1 (heatmap)

```{r ms_clus2_5}
msphate_new_idents <- c('1', '2', '3', '1', '4', '5', '6')
names(msphate_new_idents) <- levels(clus)
clus <- RenameIdents(clus, msphate_new_idents)
clus$msphate_clusters <- Idents(clus)
DimPlot(clus, reduction = 'msphate', pt.size = 1)

msphate_varmarkers_clus2_2 <- FindAllMarkers(clus, features = VariableFeatures(clus))
msphate_varmarkers_clus2_2 %>%
  group_by(cluster) %>%
  top_n(n = 7, wt = avg_log2FC) -> top_de
DoHeatmap(clus, features = top_de$gene) + NoLegend()
```
- So what's with these clusters? Are they early/late stages? Are they cycling/non-cycling cells?

```{r ms_clus2_play1}
mjt <- DimPlot(clus, reduction = 'msphate', group.by = 'jansky_idents', split.by = 'timepoint2', pt.size = 1, combine = FALSE)

mt <- DimPlot(clus, reduction = 'msphate', split.by = 'timepoint2', pt.size = 1, combine=FALSE)

fp <- FeaturePlot(clus, reduction = 'msphate', split.by = 'timepoint2', features = 'CC.Difference', cols = c('red', 'green'), pt.size = 1.5, combine = FALSE)
fp[[4]] + coord_cartesian(xlim = c(-0.015, 0.02), ylim = c(-0.02, 0.02)) + NoLegend() + fp[[5]] + coord_cartesian(xlim = c(-0.015, 0.02), ylim = c(-0.02, 0.02)) + NoLegend() + fp[[6]] + coord_cartesian(xlim = c(-0.015, 0.02), ylim = c(-0.02, 0.02)) + NoLegend() + fp[[1]] + coord_cartesian(xlim = c(-0.015, 0.02), ylim = c(-0.02, 0.02)) + NoLegend() + fp[[2]] + coord_cartesian(xlim = c(-0.015, 0.02), ylim = c(-0.02, 0.02)) + NoLegend() + fp[[3]] + coord_cartesian(xlim = c(-0.015, 0.02), ylim = c(-0.02, 0.02))
```

- We see that sub-clusters 1, 2, 3 are being replaced with 4 and 6 (possibly 5)
- From Jansky annotation, we can say that 1 - connecting Chromaffins, 2 - Neuroblasts, 3 - cycling Neuroblasts, 4 - Chromaffin cells, 6 - late Chromaffin cells
- 5 is indeterminate, perhaps a novel, distinct population of cycling Neuroblasts, most prominent at 11 pcw?
- Let's look at the difference between sub-clusters 3 and 5
- I think it's possible to put these finer clusters into a global diagram, maybe just by combining ggplot objects?? Or through metadata

##### Novel subpopulation of cycling neuroblasts?

- First approach is to compare difference in markers between 3 and 5
```{r ms_clus2_play2}
diff_5_3_markers <- FindMarkers(clus, ident.1 = 5, ident.2 = 3)
diff_5_3_markers['pct_diff'] <- diff_5_3_markers['pct.1'] - diff_5_3_markers['pct.2']
diff_5_3_markers <- diff_5_3_markers[order(diff_5_3_markers$avg_log2FC), ]
diff_5_3_markers

diff_5_all_markers <- FindMarkers(clus, ident.1 = 5)
diff_5_all_markers['pct_diff'] <- diff_5_all_markers['pct.1'] - diff_5_all_markers['pct.2']
diff_5_all_markers <- diff_5_all_markers[order(diff_5_all_markers$avg_log2FC), ]
diff_5_all_markers

diff_53_all_markers <- FindMarkers(clus, cells.1 = WhichCells()) # Only FGF13

unique_genes <- setdiff(diff_5_3_markers[1, ], diff_5_all_markers[1, ]) # Nothing, because FGF13 is turned off in cluster 5
unique_genes
```

- It seems that the absence of FGF13 is the only distinction between clusters 5 and 3
- Second approach is to take a subset of clus that includes 3 and 5, and zoom into it (same process as above)

#### Using Louvain
- We could also find sub-clusters of global msPHATE cluster 2 using Seurat

``` {r seur_clus2}
clus <- ScaleData(clus)
clus <- RunPCA(object = clus)
clus <- FindNeighbors(object = clus)
clus <- FindClusters(object = clus, resolution = 0.3)
clus <- RunUMAP(object = clus, features = VariableFeatures(clus))
DimPlot(object = clus, reduction = "umap", pt.size=1)
```
