---
title: "R Notebook"
output: html_notebook
---
## msPHATE dimensionality reduction with 2k genes
### Restart
```{r restart}
library(dplyr)
library(Seurat)
library(patchwork)
library(Matrix)
library(ggplot2)
library(tibble)
library(cowplot)
library(purrr)

wdir <- '~/Jansky/ms_phate/msphate_jansky/msPHATE-Seurat/'
clus0_2_fix_2k <- readRDS(paste(wdir, 'clus0_2_fix_2k.RDS', sep = ''))
```

### Prework
```{r prework}
library(dplyr)
library(Seurat)
library(patchwork)
library(Matrix)
library(ggplot2)
library(tibble)

wdir <- '~/Jansky/ms_phate/msphate_jansky/msPHATE-Seurat/'
data_name <- 'adrenal_medulla_Seurat.RDS.gz'
jansky <- readRDS(paste(wdir, data_name, sep = ''))
med_2k <- jansky

med_2k <- FindVariableFeatures(med_2k)
top_varfeats_global_2k <- head(VariableFeatures(med_2k), 10)
LabelPoints(VariableFeaturePlot(med_2k), points = top_varfeats_global_2k, repel = TRUE)

GetAssayData(med_2k, 'scale.data') %>% as.data.frame() %>% rownames_to_column(var='genes') %>% filter(genes %in% VariableFeatures(med_2k)) %>% column_to_rownames(var="genes") -> scale_data_2k
write.table(scale_data_2k, file = paste(wdir, 'counts.csv', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(colnames(scale_data_2k), file = paste(wdir, 'cell_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(rownames(scale_data_2k), file = paste(wdir, 'gene_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
```

Python happens here

### Choosing clustering level
```{r embedding}
msPHATE_embedding_global_2k <- read.csv(paste(wdir, 'msPHATE_embedding.csv', sep = ''), header=FALSE, row.names = 1, col.names = c('', 'msPHATE_1', 'msPHATE_2'))
msPHATE_embedding_global_2k <- as.matrix(msPHATE_embedding_global_2k)
med_2k[['msphate']] <- CreateDimReducObject(embeddings = msPHATE_embedding_global_2k, key = 'msPHATE_')
```

```{r clusters}
msPHATE_clusters_global_lvl1_2k <- as.matrix(read.csv(paste(wdir, 'msPHATE_clusters.csv', sep = ''), header=FALSE))
msPHATE_clusters_global_lvl1_2k <- factor(msPHATE_clusters_global_lvl1_2k)
med_2k$msphate_clusters_lvl1 <- msPHATE_clusters_global_lvl1_2k

msPHATE_clusters_global_lvl2_2k <- as.matrix(read.csv(paste(wdir, 'msPHATE_clusters.csv', sep = ''), header=FALSE))
msPHATE_clusters_global_lvl2_2k <- factor(msPHATE_clusters_global_lvl2_2k)
med_2k$msphate_clusters_lvl2 <- msPHATE_clusters_global_lvl2_2k

msPHATE_clusters_global_lvl3_2k <- as.matrix(read.csv(paste(wdir, 'msPHATE_clusters.csv', sep = ''), header=FALSE))
msPHATE_clusters_global_lvl3_2k <- factor(msPHATE_clusters_global_lvl3_2k)
med_2k$msphate_clusters_lvl3 <- msPHATE_clusters_global_lvl3_2k

msPHATE_clusters_global_lvl4_2k <- as.matrix(read.csv(paste(wdir, 'msPHATE_clusters.csv', sep = ''), header=FALSE))
msPHATE_clusters_global_lvl4_2k <- factor(msPHATE_clusters_global_lvl4_2k)
med_2k$msphate_clusters_lvl4 <- msPHATE_clusters_global_lvl4_2k

msPHATE_clusters_global_lvl5_2k <- as.matrix(read.csv(paste(wdir, 'msPHATE_clusters.csv', sep = ''), header=FALSE))
msPHATE_clusters_global_lvl5_2k <- factor(msPHATE_clusters_global_lvl5_2k)
med_2k$msphate_clusters_lvl5 <- msPHATE_clusters_global_lvl5_2k

msPHATE_clusters_global_lvl6_2k <- as.matrix(read.csv(paste(wdir, 'msPHATE_clusters.csv', sep = ''), header=FALSE))
msPHATE_clusters_global_lvl6_2k <- factor(msPHATE_clusters_global_lvl6_2k)
med_2k$msphate_clusters_lvl6 <- msPHATE_clusters_global_lvl6_2k

msPHATE_clusters_global_lvl7_2k <- as.matrix(read.csv(paste(wdir, 'msPHATE_clusters.csv', sep = ''), header=FALSE))
msPHATE_clusters_global_lvl7_2k <- factor(msPHATE_clusters_global_lvl7_2k)
med_2k$msphate_clusters_lvl7 <- msPHATE_clusters_global_lvl7_2k

msPHATE_clusters_global_lvl8_2k <- as.matrix(read.csv(paste(wdir, 'msPHATE_clusters.csv', sep = ''), header=FALSE))
msPHATE_clusters_global_lvl8_2k <- factor(msPHATE_clusters_global_lvl8_2k)
med_2k$msphate_clusters_lvl8 <- msPHATE_clusters_global_lvl8_2k

msPHATE_clusters_global_lvl9_2k <- as.matrix(read.csv(paste(wdir, 'msPHATE_clusters.csv', sep = ''), header=FALSE))
msPHATE_clusters_global_lvl9_2k <- factor(msPHATE_clusters_global_lvl9_2k)
med_2k$msphate_clusters_lvl9 <- msPHATE_clusters_global_lvl9_2k

msPHATE_clusters_global_lvl10_2k <- as.matrix(read.csv(paste(wdir, 'msPHATE_clusters.csv', sep = ''), header=FALSE))
msPHATE_clusters_global_lvl10_2k <- factor(msPHATE_clusters_global_lvl10_2k)
med_2k$msphate_clusters_lvl10 <- msPHATE_clusters_global_lvl10_2k
```

```{r plots}
med_2k$jansky_idents <- Idents(jansky)

DimPlot(med_2k, reduction = 'msphate', group.by = 'msphate_clusters_lvl2', pt.size = 1) + DimPlot(med_2k, reduction = 'msphate', group.by = 'msphate_clusters_lvl3', pt.size = 1) + DimPlot(med_2k, reduction = 'msphate', group.by = 'msphate_clusters_lvl4', pt.size = 1) + DimPlot(med_2k, reduction = 'msphate', group.by = 'msphate_clusters_lvl5', pt.size = 1)
DimPlot(med_2k, reduction = 'msphate', group.by = 'jansky_idents', pt.size = 1)

DimPlot(med_2k, reduction = 'msphate', group.by = 'msphate_clusters_lvl1', pt.size = 1) + DimPlot(med_2k, reduction = 'msphate', group.by = 'msphate_clusters_lvl6', pt.size = 1) + DimPlot(med_2k, reduction = 'msphate', group.by = 'msphate_clusters_lvl7', pt.size = 1) + DimPlot(med_2k, reduction = 'msphate', group.by = 'msphate_clusters_lvl8', pt.size = 1) + DimPlot(med_2k, reduction = 'msphate', group.by = 'msphate_clusters_lvl9', pt.size = 1) + DimPlot(med_2k, reduction = 'msphate', group.by = 'msphate_clusters_lvl10', pt.size = 1)
```

```{r heatmpas}
Idents(med_2k) <- 'msphate_clusters_lvl2'
global_markers_lvl2_2k <- FindAllMarkers(med_2k, features = VariableFeatures(med_2k))
global_markers_lvl2_2k %>%
  group_by(cluster) %>%
  top_n(n = 10, wt = avg_log2FC) -> top10_global_markers_lvl2_2k

Idents(med_2k) <- 'msphate_clusters_lvl3'
global_markers_lvl3_2k <- FindAllMarkers(med_2k, features = VariableFeatures(med_2k))
global_markers_lvl3_2k %>%
  group_by(cluster) %>%
  top_n(n = 10, wt = avg_log2FC) -> top10_global_markers_lvl3_2k
DoHeatmap(med_2k, features = top10_global_markers_lvl3_2k$gene) + NoLegend()

Idents(med_2k) <- 'msphate_clusters_lvl4'
global_markers_lvl4_2k <- FindAllMarkers(med_2k, features = VariableFeatures(med_2k))
global_markers_lvl4_2k %>%
  group_by(cluster) %>%
  top_n(n = 10, wt = avg_log2FC) -> top10_global_markers_lvl4_2k
DoHeatmap(med_2k, features = top10_global_markers_lvl4_2k$gene) + NoLegend()

Idents(med_2k) <- 'msphate_clusters_lvl5'
global_markers_lvl5_2k <- FindAllMarkers(med_2k, features = VariableFeatures(med_2k))
global_markers_lvl5_2k %>%
  group_by(cluster) %>%
  top_n(n = 10, wt = avg_log2FC) -> top10_global_markers_lvl5_2k
DoHeatmap(med_2k, features = top10_global_markers_lvl5_2k$gene) + NoLegend()
```

(lvl4)1 and 5472 beg merging, when zooming in later I'll keep them together. From lvl5 heatmap, I can see they are different in some important way, and I'll see that by zooming in
lvl4 best corresponds to jansky: 0 - bridge, 1+2736+5472 - chromaffins (connecting and proper), 2 - neuroblasts (proper, cycling, late - resolves in lvl5, can try to confirm by zooming), 10+3784 - SCPs

Zoom into (lvl4)1+2736+5472 and (lvl2)0 and (lvl4)2 to resolve neuroblasts and chromaffins, (lvl4)10+3784 and (lvl3)0 to resolve bridge and SCPs


## Resolving bridge and SCPs
Zooming into cluster 0(lvl2)
```{r b+s prework}
clus0_2_2k <- subset(med_2k, subset = msphate_clusters_lvl2 %in% c('0')) # Syntax allows multiple clusters
scale_data_clus0_2_2k <- GetAssayData(clus0_2_2k, 'scale.data')
write.table(scale_data_clus0_2_2k, 
            file = paste(wdir, 'counts.csv', sep = ''), 
            sep = ',', row.names = FALSE, col.names = FALSE)
write.table(rownames(scale_data_clus0_2_2k), file = paste(wdir, 'gene_names.txt', sep = ''), 
            sep = ',', row.names = FALSE, col.names = FALSE)
write.table(colnames(scale_data_clus0_2_2k), file = paste(wdir, 'cell_names.txt', sep = ''), 
            sep = ',', row.names = FALSE, col.names = FALSE)
```

```{r b+s embedding}
msPHATE_embedding_clus0_2_2k <- read.csv(paste(wdir, 'msPHATE_embedding.csv', sep = ''), header=FALSE, row.names = 1, col.names = c('', 'msPHATE_1', 'msPHATE_2'))
msPHATE_embedding_clus0_2_2k <- as.matrix(msPHATE_embedding_clus0_2_2k)
clus0_2_2k[['msphate']] <- CreateDimReducObject(embeddings = msPHATE_embedding_clus0_2_2k, key = 'msPHATE_')
```

```{r b+s clusters}
msPHATE_clusters_clus0_2_2k <- as.matrix(read.csv(paste(wdir, 'msPHATE_clusters.csv', sep = ''), header=FALSE))
msPHATE_clusters_clus0_2_2k <- factor(msPHATE_clusters_clus0_2_2k)
clus0_2_2k$msphate_subclusters_lvl3 <- msPHATE_clusters_clus0_2_2k
```

```{r b+s plot}
DimPlot(clus0_2_2k, reduction = 'msphate', group.by = 'msphate_subclusters_lvl3', pt.size = 1) + DimPlot(clus0_2_2k, reduction = 'msphate', group.by = 'jansky_idents', pt.size = 1)
```

This is a bit weird, definitely need to get rid of subcluster 237

```{r b+s fix}
clus_0_2_oddcells_2k <- WhichCells(clus0_2_2k, expression = msphate_subclusters_lvl3 %in% c('237') | jansky_idents %in% c('connecting Chromaffin cells', 'Chromaffin cells', 'late Chromaffin cells', 'cycling Neuroblasts', 'Neuroblasts', 'late Neuroblasts'))
length(clus_0_2_oddcells_2k) # 62 cells - likely noise

clus0_2_fix_2k <- subset(clus0_2_2k, subset = msphate_subclusters_lvl3 %in% c('0', '1', '19'))
clus0_2_fix_2k <- subset(clus0_2_fix_2k, subset = jansky_idents %in% c('late SCPs', 'SCPs', 'cycling SCPs', 'Bridge'))
DimPlot(clus0_2_fix_2k, reduction = 'msphate', group.by = 'msphate_subclusters_lvl3', pt.size = 1) + DimPlot(clus0_2_fix_2k, reduction = 'msphate', group.by = 'jansky_idents', pt.size = 1)
```

Now need to recluster and different resolutions (clustering will change even at lvl3)

```{r clus0_2_fix_lvl4}
scale_data_clus0_2_fix_2k <- GetAssayData(clus0_2_fix_2k, 'scale.data')
write.table(scale_data_clus0_2_fix_2k, 
            file = paste(wdir, 'counts.csv', sep = ''), 
            sep = ',', row.names = FALSE, col.names = FALSE)
write.table(rownames(scale_data_clus0_2_fix_2k), file = paste(wdir, 'gene_names.txt', sep = ''), 
            sep = ',', row.names = FALSE, col.names = FALSE)
write.table(colnames(scale_data_clus0_2_fix_2k), file = paste(wdir, 'cell_names.txt', sep = ''), 
            sep = ',', row.names = FALSE, col.names = FALSE)

msPHATE_embedding_clus0_2_fix_2k <- read.csv(paste(wdir, 'msPHATE_embedding.csv', sep = ''), header=FALSE, row.names = 1, col.names = c('', 'msPHATE_1', 'msPHATE_2'))
msPHATE_embedding_clus0_2_fix_2k <- as.matrix(msPHATE_embedding_clus0_2_fix_2k)
clus0_2_fix_2k[['msphate']] <- CreateDimReducObject(embeddings = msPHATE_embedding_clus0_2_fix_2k, key = 'msPHATE_')

msPHATE_clusters_clus0_2_fix_2k <- as.matrix(read.csv(paste(wdir, 'msPHATE_clusters.csv', sep = ''), header=FALSE))
msPHATE_clusters_clus0_2_fix_2k <- factor(msPHATE_clusters_clus0_2_fix_2k)
clus0_2_fix_2k$msphate_subclusters_lvl4 <- msPHATE_clusters_clus0_2_fix_2k

msPHATE_clusters_clus0_2_fix_2k <- as.matrix(read.csv(paste(wdir, 'msPHATE_clusters.csv', sep = ''), header=FALSE))
msPHATE_clusters_clus0_2_fix_2k <- factor(msPHATE_clusters_clus0_2_fix_2k)
clus0_2_fix_2k$msphate_subclusters_lvl3 <- msPHATE_clusters_clus0_2_fix_2k

msPHATE_clusters_clus0_2_fix_2k <- as.matrix(read.csv(paste(wdir, 'msPHATE_clusters.csv', sep = ''), header=FALSE))
msPHATE_clusters_clus0_2_fix_2k <- factor(msPHATE_clusters_clus0_2_fix_2k)
clus0_2_fix_2k$msphate_subclusters_lvl2 <- msPHATE_clusters_clus0_2_fix_2k

DimPlot(clus0_2_fix_2k, reduction = 'msphate', group.by = 'msphate_subclusters_lvl2', pt.size = 1) + DimPlot(clus0_2_fix_2k, reduction = 'msphate', group.by = 'msphate_subclusters_lvl3', pt.size = 1) + DimPlot(clus0_2_fix_2k, reduction = 'msphate', group.by = 'msphate_subclusters_lvl4', pt.size = 1) + DimPlot(clus0_2_fix_2k, reduction = 'msphate', group.by = 'jansky_idents', pt.size = 1)
```

```{r clus0_2_fix heatmpas}
clus0_2_fix_2k <- FindVariableFeatures(clus0_2_fix_2k)
top_varfeats_clus0_2_fix_2k <- head(VariableFeatures(clus0_2_fix_2k), 10)
LabelPoints(VariableFeaturePlot(clus0_2_fix_2k), points = top_varfeats_clus0_2_fix_2k, repel = TRUE)

Idents(clus0_2_fix_2k) <- 'msphate_subclusters_lvl3'
clus0_2_fix_markers_lvl3_2k <- FindAllMarkers(clus0_2_fix_2k, features = VariableFeatures(clus0_2_fix_2k))
clus0_2_fix_markers_lvl3_2k %>%
  group_by(cluster) %>%
  top_n(n = 10, wt = avg_log2FC) -> top10_clus0_2_fix_markers_lvl3_2k
DoHeatmap(clus0_2_fix_2k, features = top10_clus0_2_fix_markers_lvl3_2k$gene) + NoLegend()
```

Once again, (lvl3)1 and 1185 beg to be merged. Once again, one is a low number and the other is hihj. We saw a similar thing with global (lvl3)1 and 5472. 

```{r diff_1185_1_clus0_2_fix_2k}
Idents(clus0_2_fix_2k) <- 'msphate_subclusters_lvl3'
diff_1185_1_clus0_2_fix_2k <- FindMarkers(clus0_2_fix_2k, ident.1 = 1185, ident.2 = 1)
diff_1185_1_clus0_2_fix_2k['pct_diff'] <- diff_1185_1_clus0_2_fix_2k['pct.1'] - diff_1185_1_clus0_2_fix_2k['pct.2']
diff_1185_1_clus0_2_fix_2k <- diff_1185_1_clus0_2_fix_2k[order(diff_1185_1_clus0_2_fix_2k$avg_log2FC), ]

diff_1185_all_clus0_2_fix_2k <- FindMarkers(clus0_2_fix_2k, ident.1 = 1185)
diff_1185_all_clus0_2_fix_2k['pct_diff'] <- diff_1185_all_clus0_2_fix_2k['pct.1'] - diff_1185_all_clus0_2_fix_2k['pct.2']
diff_1185_all_clus0_2_fix_2k <- diff_1185_all_clus0_2_fix_2k[order(diff_1185_all_clus0_2_fix_2k$avg_log2FC), ]

diff_1_all_clus0_2_fix_2k <- FindMarkers(clus0_2_fix_2k, ident.1 = 1)
diff_1_all_clus0_2_fix_2k['pct_diff'] <- diff_1_all_clus0_2_fix_2k['pct.1'] - diff_1_all_clus0_2_fix_2k['pct.2']
diff_1_all_clus0_2_fix_2k <- diff_1_all_clus0_2_fix_2k[order(diff_1_all_clus0_2_fix_2k$avg_log2FC), ]

diff_1185_1_clus0_2_fix_2k
diff_1185_all_clus0_2_fix_2k
diff_1_all_clus0_2_fix_2k

d_1185_1__1185_all <- setdiff(diff_1185_1_clus0_2_fix_2k[1, ], diff_1185_all_clus0_2_fix_2k[1, ])
d_1185_all__1185_1 <- setdiff(diff_1185_all_clus0_2_fix_2k[1, ], diff_1185_1_clus0_2_fix_2k[1, ])

d_1185_1__1_all <- setdiff(diff_1185_1_clus0_2_fix_2k[1, ], diff_1_all_clus0_2_fix_2k[1, ])
d_1_all__1185_1 <- setdiff(diff_1_all_clus0_2_fix_2k[1, ], diff_1185_1_clus0_2_fix_2k[1, ])

d_1185_all__1_all <- setdiff(diff_1185_all_clus0_2_fix_2k[1, ], diff_1_all_clus0_2_fix_2k[1, ])
d_1_all__1185_all <- setdiff(diff_1_all_clus0_2_fix_2k[1, ], diff_1185_all_clus0_2_fix_2k[1, ])

d_1185_1__1185_all # PDZRN4
d_1185_all__1185_1 # PDZRN4
d_1185_1__1_all # PDZRN4
d_1_all__1185_1 #VWDE
d_1185_all__1_all # PDZRN4
d_1_all__1185_all #VWDE

pdzrn_featplot_clus0_2_fix_2k <- FeaturePlot(clus0_2_fix_2k, reduction = 'msphate', features = 'PDZRN4', pt.size = 1.1) + coord_cartesian(xlim = c(-0.025, 0.065), ylim = c(-0.035, 0.035))
vwde_featplot_clus0_2_fix_2k <- FeaturePlot(clus0_2_fix_2k, reduction = 'msphate', features = 'VWDE', pt.size = 1.1) + coord_cartesian(xlim = c(-0.025, 0.065), ylim = c(-0.035, 0.035))
ccdiff_featplot_clus2_0_fix_2k <- FeaturePlot(clus0_2_fix_2k, reduction = 'msphate', features = 'CC.Difference', cols = c('red', 'green'), pt.size = 1.1) + coord_cartesian(xlim = c(-0.025, 0.065), ylim = c(-0.035, 0.035))

pdzrn_featplot_clus0_2_fix_2k + vwde_featplot_clus0_2_fix_2k + ccdiff_featplot_clus2_0_fix_2k +  DimPlot(clus0_2_fix_2k, reduction = 'msphate', group.by = 'msphate_subclusters_lvl3', pt.size = 1) + DimPlot(clus0_2_fix_2k, reduction = 'msphate', group.by = 'jansky_idents', pt.size = 1)

DimPlot(clus0_2_fix_2k, reduction = 'msphate', split.by = 'timepoint2', group.by = 'msphate_subclusters_lvl3', pt.size = 1)
```

Need somebody to check my thinking
Next, I found that it's possible to essentially FeaturePlot PCs driving cluters

```{r clus0_2_fix_2k PCs}
pc1 <- FeaturePlot(clus0_2_fix_2k, features = "PC_1", reduction = 'msphate') + coord_cartesian(xlim = c(-0.025, 0.065), ylim = c(-0.035, 0.035))
pc2 <- FeaturePlot(clus0_2_fix_2k, features = "PC_2", reduction = 'msphate') + coord_cartesian(xlim = c(-0.025, 0.065), ylim = c(-0.035, 0.035))
pc3 <- FeaturePlot(clus0_2_fix_2k, features = "PC_3", reduction = 'msphate') + coord_cartesian(xlim = c(-0.025, 0.065), ylim = c(-0.035, 0.035))
pc1+pc2+pc3

# pc3 - pc5 + pc4
```

```{r pc math}
pcs <- FetchData(clus0_2_fix_2k, vars = c('PC_3', 'PC_5', 'PC_4', 'msPHATE_1', 'msPHATE_2'))
pcs['score'] <- pcs['PC_3']-pcs['PC_5']+pcs['PC_4']

loadings <- clus0_2_fix_2k@reductions$pca@feature.loadings
clus0_2_fix_2k$score_1185 <- loadings[, 'PC_3'] - loadings[, 'PC_5'] + loadings[, 'PC_4']
FeaturePlot(clus0_2_fix_2k, features = "score_1185", reduction = 'msphate') + coord_cartesian(xlim = c(-0.025, 0.065), ylim = c(-0.035, 0.035))
```
