---
title: "R Notebook"
output: html_notebook
---

```{r}
## RESTART ##
library(dplyr)
library(Seurat)
library(patchwork)
library(Matrix)
library(ggplot2)
library(tibble)
library(cowplot)
library(purrr)
wdir <- '~/Jansky/ms_phate/msphate_jansky/msPHATE-Seurat/'
data_name <- 'med_final.RDS'
med <- readRDS(paste(wdir, data_name, sep = ''))
data_name <- 'bcn_final.RDS'
bcn <- readRDS(paste(wdir, data_name, sep = ''))
data_name <- 'bcn_ms_final.RDS'
bcn_ms <- readRDS(paste(wdir, data_name, sep = ''))
```

## 1. Top 2k most variable genes vs all genes
```{r}
library(dplyr)
library(Seurat)
library(patchwork)
library(Matrix)
library(ggplot2)
library(tibble)
library(cowplot)
library(purrr)
wdir <- '~/Jansky/ms_phate/msphate_jansky/msPHATE-Seurat/'

data_name <- 'adrenal_medulla_Seurat.RDS.gz'
med <- readRDS(paste(wdir, data_name, sep = ''))
```

```{r}
top_varfeats_global <- head(VariableFeatures(med), 10)
LabelPoints(VariableFeaturePlot(med), points = top_varfeats_global, repel = TRUE)
```

Now we'll cluster based only on expression of the most variable genes (initial seurat object unchanged, embedding and cluster identity are metadata belonging to cells, and every gene is still measured in every cell, so subsetting by genes still keeps all cells). These are the same genes used in the Jansky analysis.

Python code runs well in R, but faster to run msphate.py externally

```{python}
### USAGE ###
# Export counts.csv. gene_names.txt and cell.names.txt from R
# Run this script (F5)
# Shell restarts, files loaded in, msPHATE tree constructed
# Now you can run msPHATE on the same data multiple times at different clustering levels
# Have to restart again to change other parameters as they are hard-coded at the moment
# To use newly written data, run the script again (shell restarts), then msphate(clustering_level)
### USAGE ###

import multiscale_phate as mp
import numpy as np
import pandas as pd
import scprep
import os

wdir = '~/Jansky/ms_phate/msphate_jansky/msPHATE-Seurat'
npca = 20 # Deafault value for RunPCA in Seurat
gran = 0.1
vis_level = 0 # Always zero so points on embedding correspond to individual cells

wdir = os.path.expanduser(wdir)
gene_path = os.path.join(wdir, 'gene_names.txt')
cell_path = os.path.join(wdir, 'cell_names.txt')

print('Loading scaled data...')
data_path = os.path.join(wdir, 'counts.csv')
data = scprep.io.load_csv(data_path, cell_axis='column',
                        gene_names=gene_path,
                        cell_names=cell_path)

mp_op = mp.Multiscale_PHATE(n_pca=npca, granularity=gran, random_state=0)
levels = mp_op.fit(data)

def msphate(clustering_level):
    clus_level = clustering_level
    print('Generating embedding and assigning clusters...')
    embedding, clusters, sizes = mp_op.transform(visualization_level = levels[vis_level],
                                                               cluster_level = levels[-1*clus_level])

    print('Exporting embedding and clusters...')
    msPHATE_embedding = pd.DataFrame(embedding,
                          index = pd.read_csv(cell_path, header=None).iloc[:, 0].to_list())
    msPHATE_embedding.to_csv('msPHATE_embedding.csv', header=False)
    msPHATE_clusters = pd.DataFrame(clusters)
    msPHATE_clusters.to_csv('msPHATE_clusters.csv', header=False, index=False)

    print('\n')
    print(data)
    print('\n')
    print(levels)
    print('\n')
    print(embedding)
    print('\n')
    print(set(clusters))
    print('Done')
```

### Top 2k genes
```{r}
FALSE %in% (rownames(GetAssayData(med, 'scale.data')) %in% VariableFeatures(med)) # Any difference between varfeats and scale.data?
FALSE %in% (rownames(scale_data_2k) %in% rownames(GetAssayData(med, 'scale.data'))) # Any difference between 'extracting' top 2k genes from scale.data and scale.data itself?
# No to both
# Means scale.data is already for top 2k genes
```

```{r}
write.table(GetAssayData(med, 'scale.data'), file = paste(wdir, 'counts.csv', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(colnames(GetAssayData(med, 'scale.data')), file = paste(wdir, 'cell_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(rownames(GetAssayData(med, 'scale.data')), file = paste(wdir, 'gene_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
# Files written are run through msPHATE in Python
```

```{python}
msphate(3)
# Use msphate.py
```

```{r}
msPHATE_embedding_global <- read.csv(paste(wdir, 'msPHATE_embedding.csv', sep = ''), header=FALSE, row.names = 1, col.names = c('', 'msPHATE_1', 'msPHATE_2'))
msPHATE_embedding_global <- as.matrix(msPHATE_embedding_global)
med[['msphate_2k']] <- CreateDimReducObject(embeddings = msPHATE_embedding_global, key = 'msPHATE_')
# Only run once - embedding constant for different clustering levels
```

```{r}
msPHATE_clusters_global <- as.matrix(read.csv(paste(wdir, 'msPHATE_clusters.csv', sep = ''), header=FALSE))
msPHATE_clusters_global <- factor(msPHATE_clusters_global)
med$msphate_clusters_lvl3_2k <- msPHATE_clusters_global
```

```{python}
msphate(4)
# Use msphate.py
```

```{r}
msPHATE_clusters_global <- as.matrix(read.csv(paste(wdir, 'msPHATE_clusters.csv', sep = ''), header=FALSE))
msPHATE_clusters_global <- factor(msPHATE_clusters_global)
med$msphate_clusters_lvl4_2k <- msPHATE_clusters_global
```

```{python}
msphate(5)
# Use msphate.py
```

```{r}
msPHATE_clusters_global <- as.matrix(read.csv(paste(wdir, 'msPHATE_clusters.csv', sep = ''), header=FALSE))
msPHATE_clusters_global <- factor(msPHATE_clusters_global)
med$msphate_clusters_lvl5_2k <- msPHATE_clusters_global
```

```{python}
msphate(2)
# Use msphate.py
```

```{r}
msPHATE_clusters_global <- as.matrix(read.csv(paste(wdir, 'msPHATE_clusters.csv', sep = ''), header=FALSE))
msPHATE_clusters_global <- factor(msPHATE_clusters_global)
med$msphate_clusters_lvl2_2k <- msPHATE_clusters_global
```

### All genes
```{r}
med_all <- med
med_all <- NormalizeData(med_all)
med_all <- ScaleData(object = med_all, features = rownames(med_all@assays$RNA@counts))
Misc(med, 'scale.data_all') <- GetAssayData(med_all, 'scale.data')
med@misc$scale.data_all[1:5, 1:5]
rm(med_all)

write.table(med@misc$scale.data_all, file = paste(wdir, 'counts.csv', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(colnames(med@misc$scale.data_all), file = paste(wdir, 'cell_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(rownames(med@misc$scale.data_all), file = paste(wdir, 'gene_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
```

```{python}
msphate(3)
# Use msphate.py
```

```{r}
msPHATE_embedding_global <- read.csv(paste(wdir, 'msPHATE_embedding.csv', sep = ''), header=FALSE, row.names = 1, col.names = c('', 'msPHATE_1', 'msPHATE_2'))
msPHATE_embedding_global <- as.matrix(msPHATE_embedding_global)
med[['msphate_all']] <- CreateDimReducObject(embeddings = msPHATE_embedding_global, key = 'msPHATE_')
```

```{r}
msPHATE_clusters_global <- as.matrix(read.csv(paste(wdir, 'msPHATE_clusters.csv', sep = ''), header=FALSE))
msPHATE_clusters_global <- factor(msPHATE_clusters_global)
med$msphate_clusters_lvl3_all <- msPHATE_clusters_global
```

```{python}
msphate(4)
# Use msphate.py
```

```{r}
msPHATE_clusters_global <- as.matrix(read.csv(paste(wdir, 'msPHATE_clusters.csv', sep = ''), header=FALSE))
msPHATE_clusters_global <- factor(msPHATE_clusters_global)
med$msphate_clusters_lvl4_all <- msPHATE_clusters_global
```

```{python}
msphate(5)
# Use msphate.py
```

```{r}
msPHATE_clusters_global <- as.matrix(read.csv(paste(wdir, 'msPHATE_clusters.csv', sep = ''), header=FALSE))
msPHATE_clusters_global <- factor(msPHATE_clusters_global)
med$msphate_clusters_lvl5_all <- msPHATE_clusters_global
```

```{python}
msphate(2)
# Use msphate.py
```

```{r}
msPHATE_clusters_global <- as.matrix(read.csv(paste(wdir, 'msPHATE_clusters.csv', sep = ''), header=FALSE))
msPHATE_clusters_global <- factor(msPHATE_clusters_global)
med$msphate_clusters_lvl2_all <- msPHATE_clusters_global
```

Let's visualise the clustering and see how using the top 2k most variable genes vs all the genes affects msPHATE clustering compared to UMAP

```{r}
med[['pca_all']] <- CreateDimReducObject(embeddings = Embeddings(med[['pca']]), key = 'PC_')
med@reductions$pca <- NULL

umap_tsne_all <- med
med$jansky_idents <- Idents(med)
all_genes <- rownames(GetAssayData(umap_tsne_all, 'scale.data'))
umap_tsne_all <- ScaleData(umap_tsne_all, features = all_genes)
umap_tsne_all <- RunPCA(object = umap_tsne_all, features = all_genes, reduction.name = 'pca_all')
umap_tsne_all <- FindNeighbors(object = umap_tsne_all, reduction = 'pca_all')
umap_tsne_all <- FindClusters(object = umap_tsne_all, resolution = 1.2)
umap_tsne_all <- RunUMAP(object = umap_tsne_all, features = all_genes)
umap_tsne_all <- RunTSNE(object = umap_tsne_all, features = all_genes)
med <- RunTSNE(object = med, features = VariableFeatures(med), reduction = 'pca_all')
rm(umap_tsne_all)

med[['umap_2k']] <- CreateDimReducObject(embeddings = Embeddings(med[['umap']]), key = 'UMAP_')
med[['tsne_2k']] <- CreateDimReducObject(embeddings = Embeddings(med[['tsne']]), key = 'UMAP_')
med@reductions$umap <- NULL
med@reductions$tsne <- NULL

med[['umap_all']] <- CreateDimReducObject(embeddings = Embeddings(umap_tsne_all[['umap']]), key = 'UMAP_')
med[['tsne_all']] <- CreateDimReducObject(embeddings = Embeddings(umap_tsne_all[['tsne']]), key = 'UMAP_')
```

```{r}
DimPlot(med, reduction = 'msphate_2k', group.by = 'jansky_idents', pt.size = 1) + ggtitle('msphate_2k by jansky_idents')
DimPlot(med, reduction = 'msphate_all', group.by = 'jansky_idents', pt.size = 1) + ggtitle('msphate_all by jansky_idents')

DimPlot(med, reduction = 'msphate_2k', group.by = 'msphate_clusters_lvl2_2k', pt.size = 1) + ggtitle('msphate_2k at lvl2')
DimPlot(med, reduction = 'msphate_2k', group.by = 'msphate_clusters_lvl3_2k', pt.size = 1) + ggtitle('msphate_2k at lvl3')
DimPlot(med, reduction = 'msphate_2k', group.by = 'msphate_clusters_lvl4_2k', pt.size = 1) + ggtitle('msphate_2k at lvl4')
DimPlot(med, reduction = 'msphate_2k', group.by = 'msphate_clusters_lvl5_2k', pt.size = 1) + ggtitle('msphate_2k at lvl5')

DimPlot(med, reduction = 'msphate_all', group.by = 'msphate_clusters_lvl2_all', pt.size = 1) + ggtitle('msphate_all at lvl2')
DimPlot(med, reduction = 'msphate_all', group.by = 'msphate_clusters_lvl3_all', pt.size = 1) + ggtitle('msphate_all at lvl3')
DimPlot(med, reduction = 'msphate_all', group.by = 'msphate_clusters_lvl4_all', pt.size = 1) + ggtitle('msphate_all at lvl4')
DimPlot(med, reduction = 'msphate_all', group.by = 'msphate_clusters_lvl5_all', pt.size = 1) + ggtitle('msphate_all at lvl5')

DimPlot(med, reduction = 'umap_2k', group.by = 'jansky_idents', pt.size = 1) + ggtitle('umap_2k by jansky_idents')
DimPlot(med, reduction = 'umap_2k', group.by = 'seurat_clusters', pt.size = 1) + ggtitle('umap_2k by seurat_clusters')
DimPlot(med, reduction = 'umap_all', group.by = 'jansky_idents', pt.size = 1) + ggtitle('umap_all by jansky_idents')
DimPlot(med, reduction = 'umap_all', group.by = 'seurat_clusters', pt.size = 1) + ggtitle('umap_2k by seurat_clusters')

DimPlot(med, reduction = 'tsne_2k', group.by = 'jansky_idents', pt.size = 1) + ggtitle('tsne_2k by jansky_idents')
DimPlot(med, reduction = 'tsne_2k', group.by = 'seurat_clusters', pt.size = 1) + ggtitle('tsne_2k by seurat_clusters')
DimPlot(med, reduction = 'tsne_all', group.by = 'jansky_idents', pt.size = 1) + ggtitle('tsne_all by jansky_idents')
DimPlot(med, reduction = 'tsne_all', group.by = 'seurat_clusters', pt.size = 1) + ggtitle('tsne_all by seurat_clusters')
```


## 2. Scaling vs not scaling

```{r}
med_unscaled <- med
med_unscaled <- NormalizeData(med_unscaled)
med_unscaled <- ScaleData(med_unscaled, do.scale = FALSE)
Misc(med, 'scale.data_unscaled') <- GetAssayData(med_unscaled, 'scale.data')
med@misc$scale.data_unscaled[1:5, 1:5]
rm(med_unscaled)
```

```{r}
write.table(med@misc$scale.data_unscaled, file = paste(wdir, 'counts.csv', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(colnames(med@misc$scale.data_unscaled), file = paste(wdir, 'cell_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(rownames(med@misc$scale.data_unscaled), file = paste(wdir, 'gene_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
```

```{python}
msphate(3)
# Use msphate.py
```


```{r}
msPHATE_embedding_global <- read.csv(paste(wdir, 'msPHATE_embedding.csv', sep = ''), header=FALSE, row.names = 1, col.names = c('', 'msPHATE_1', 'msPHATE_2'))
msPHATE_embedding_global <- as.matrix(msPHATE_embedding_global)
med[['msphate_unscaled']] <- CreateDimReducObject(embeddings = msPHATE_embedding_global, key = 'msPHATE_')
# Only run once - embedding constant for different clustering levels
```

```{r}
msPHATE_clusters_global <- as.matrix(read.csv(paste(wdir, 'msPHATE_clusters.csv', sep = ''), header=FALSE))
msPHATE_clusters_global <- factor(msPHATE_clusters_global)
med$msphate_clusters_lvl3_unscaled <- msPHATE_clusters_global
```

```{python}
msphate(4)
# Use msphate.py
```

```{r}
msPHATE_clusters_global <- as.matrix(read.csv(paste(wdir, 'msPHATE_clusters.csv', sep = ''), header=FALSE))
msPHATE_clusters_global <- factor(msPHATE_clusters_global)
med$msphate_clusters_lvl4_unscaled <- msPHATE_clusters_global
```


```{r}
umap_tsne_unscaled <- med
umap_tsne_unscaled <- RunPCA(object = umap_tsne_unscaled, reduction.name = 'pca_unscaled')
umap_tsne_unscaled <- FindNeighbors(object = umap_tsne_unscaled, reduction = 'pca_unscaled')
umap_tsne_unscaled <- FindClusters(object = umap_tsne_unscaled, resolution = 1.2)
umap_tsne_unscaled <- RunUMAP(object = umap_tsne_unscaled, features = VariableFeatures(umap_tsne_unscaled))
umap_tsne_unscaled <- RunTSNE(object = umap_tsne_unscaled, features = VariableFeatures(umap_tsne_unscaled), reduction = 'pca_unscaled')

med[['umap_unscaled']] <- CreateDimReducObject(embeddings = Embeddings(umap_tsne_unscaled[['umap']]), key = 'UMAP_')
med[['tsne_unscaled']] <- CreateDimReducObject(embeddings = Embeddings(umap_tsne_unscaled[['tsne']]), key = 'UMAP_')
rm(umap_tsne_unscaled)
```

```{r}
DimPlot(med, reduction = 'msphate_2k', group.by = 'jansky_idents', pt.size = 1) + ggtitle('msphate_scaled by jansky_idents')
DimPlot(med, reduction = 'msphate_unscaled', group.by = 'jansky_idents', pt.size = 1) + ggtitle('msphate_unscaled by jansky_idents')
DimPlot(med, reduction = 'msphate_2k', group.by = 'msphate_clusters_lvl3_2k', pt.size = 1) + ggtitle('msphate_scaled at lvl3')
DimPlot(med, reduction = 'msphate_unscaled', group.by = 'msphate_clusters_lvl3_unscaled', pt.size = 1) + ggtitle('msphate_unscaled at lvl3')
DimPlot(med, reduction = 'msphate_2k', group.by = 'msphate_clusters_lvl4_2k', pt.size = 1) + ggtitle('msphate_scaled at lvl4')
DimPlot(med, reduction = 'msphate_2k', group.by = 'msphate_clusters_lvl4_unscaled', pt.size = 1) + ggtitle('msphate_unscaled at lvl4')

DimPlot(med, reduction = 'umap_2k', group.by = 'jansky_idents', pt.size = 1) + ggtitle('umap_scaled by jansky_idents')
DimPlot(med, reduction = 'umap_2k', group.by = 'seurat_clusters', pt.size = 1) + ggtitle('umap_scaled by seurat_clusters')
DimPlot(med, reduction = 'umap_unscaled', group.by = 'jansky_idents', pt.size = 1) + ggtitle('umap_unscaled by jansky_idents')
DimPlot(med, reduction = 'umap_unscaled', group.by = 'seurat_clusters', pt.size = 1) + ggtitle('umap_unscaled by seurat_clusters')

DimPlot(med, reduction = 'tsne_2k', group.by = 'jansky_idents', pt.size = 1) + ggtitle('tsne_scaled by jansky_idents')
DimPlot(med, reduction = 'tsne_2k', group.by = 'seurat_clusters', pt.size = 1) + ggtitle('tsne_scaled by seurat_clusters')
DimPlot(med, reduction = 'tsne_unscaled', group.by = 'jansky_idents', pt.size = 1) + ggtitle('tsne_unscaled by jansky_idents')
DimPlot(med, reduction = 'tsne_unscaled', group.by = 'seurat_clusters', pt.size = 1) + ggtitle('tsne_unscaled by seurat_clusters')
```

```{r}
#saveRDS(med, file = paste(wdir, 'med_final.RDS', sep = ''))
```


## 3. Bbridge, connecting chromaffins, and neuroblasts

### Using jansky_idents

```{r}
bcn <- subset(med, subset = jansky_idents %in% c('Bridge', 'connecting Chromaffin cells', 'Neuroblasts', 'cycling Neuroblasts', 'late Neuroblasts'))
bcn <- FindVariableFeatures(bcn)
bcn <- NormalizeData(bcn)
bcn <- ScaleData(bcn)

bcn_unscaled <- bcn
bcn_unscaled <- NormalizeData(bcn_unscaled)
bcn_unscaled <- ScaleData(bcn_unscaled, do.scale = FALSE)
Misc(bcn, 'scale.data_unscaled') <- GetAssayData(bcn_unscaled, 'scale.data')

bcn_unscaled <- RunPCA(object = bcn_unscaled)
bcn_unscaled <- FindNeighbors(object = bcn_unscaled)
bcn_unscaled <- FindClusters(object = bcn_unscaled, resolution = 1.2)
bcn_unscaled <- RunUMAP(object = bcn_unscaled, features = VariableFeatures(bcn_unscaled))
bcn_unscaled <- RunTSNE(object = bcn_unscaled, features = VariableFeatures(bcn_unscaled))
bcn[['pca_unscaled']] <- CreateDimReducObject(embeddings = Embeddings(bcn_unscaled[['pca']]), key = 'PC_')
bcn[['umap_unscaled']] <- CreateDimReducObject(embeddings = Embeddings(bcn_unscaled[['umap']]), key = 'UMAP_')
bcn[['tsne_unscaled']] <- CreateDimReducObject(embeddings = Embeddings(bcn_unscaled[['tsne']]), key = 'UMAP_')

rm(bcn_unscaled)
```

```{r}
DimPlot(med, reduction = 'msphate_2k', group.by = 'msphate_clusters_lvl2_2k', pt.size = 1)
DimPlot(med, reduction = 'msphate_2k', group.by = 'msphate_clusters_lvl3_2k', pt.size = 1)
DimPlot(med, reduction = 'msphate_2k', group.by = 'msphate_clusters_lvl4_2k', pt.size = 1)
DimPlot(med, reduction = 'msphate_2k', group.by = 'msphate_clusters_lvl5_2k', pt.size = 1)
DimPlot(med, reduction = 'msphate_2k', group.by = 'jansky_idents', pt.size = 1)
```

```{r}
DimPlot(med, reduction = 'msphate_2k', group.by = 'msphate_clusters_lvl4_2k', pt.size = 1, label = TRUE) + NoLegend()
DimPlot(med, reduction = 'msphate_2k', group.by = 'msphate_clusters_lvl5_2k', pt.size = 1, label = TRUE) + NoLegend()
DimPlot(med, reduction = 'msphate_2k', group.by = 'jansky_idents', pt.size = 1, label = TRUE) + NoLegend()
```


```{r}
# Zooming into 0, 336, 82, 1, 4607, 6, 4829, 3, 5, 2, 1814, 1838 of lvl4
bcn_ms <- subset(med, subset = msphate_clusters_lvl4_2k %in% c('0', '336', '82', '1', '4607', '6', '4829', '3', '5', '2', '1814', '1838'))
rm(med)
bcn_ms <- subset(bcn_ms, subset = jansky_idents %in% c('Bridge', 'connecting Chromaffin cells', 'Neuroblasts', 'cycling Neuroblasts', 'late Neuroblasts'))
bcn_ms <- FindVariableFeatures(bcn_ms)
bcn_ms <- NormalizeData(bcn_ms)
bcn_ms <- ScaleData(bcn_ms)

bcn_ms_unscaled <- bcn_ms
bcn_ms_unscaled <- NormalizeData(bcn_ms_unscaled)
bcn_ms_unscaled <- ScaleData(bcn_ms_unscaled, do.scale = FALSE)
Misc(bcn_ms, 'scale.data_unscaled') <- GetAssayData(bcn_ms_unscaled, 'scale.data')

bcn_ms_unscaled <- RunPCA(object = bcn_ms_unscaled)
bcn_ms_unscaled <- FindNeighbors(object = bcn_ms_unscaled)
bcn_ms_unscaled <- FindClusters(object = bcn_ms_unscaled, resolution = 1.2)
bcn_ms_unscaled <- RunUMAP(object = bcn_ms_unscaled, features = VariableFeatures(bcn_ms_unscaled))
bcn_ms_unscaled <- RunTSNE(object = bcn_ms_unscaled, features = VariableFeatures(bcn_ms_unscaled))
bcn_ms[['pca_unscaled']] <- CreateDimReducObject(embeddings = Embeddings(bcn_ms_unscaled[['pca']]), key = 'PC_')
bcn_ms[['umap_unscaled']] <- CreateDimReducObject(embeddings = Embeddings(bcn_ms_unscaled[['umap']]), key = 'UMAP_')
bcn_ms[['tsne_unscaled']] <- CreateDimReducObject(embeddings = Embeddings(bcn_ms_unscaled[['tsne']]), key = 'UMAP_')

rm(bcn_ms_unscaled)
```

```{r}
write.table(GetAssayData(bcn, 'scale.data'), file = paste(wdir, 'counts.csv', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(colnames(GetAssayData(bcn, 'scale.data')), file = paste(wdir, 'cell_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(rownames(GetAssayData(bcn, 'scale.data')), file = paste(wdir, 'gene_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
# Run msPHATE
```

```{python}
### RESTART ###
msphate(4)
# Use msphate.py
```

```{r}
msPHATE_embedding <- read.csv(paste(wdir, 'msPHATE_embedding.csv', sep = ''), header=FALSE, row.names = 1, col.names = c('', 'msPHATE_1', 'msPHATE_2'))
msPHATE_embedding <- as.matrix(msPHATE_embedding)
bcn[['msphate_scaled']] <- CreateDimReducObject(embeddings = msPHATE_embedding, key = 'msPHATE_')
```

```{r}
msPHATE_clusters <- as.matrix(read.csv(paste(wdir, 'msPHATE_clusters.csv', sep = ''), header=FALSE))
msPHATE_clusters_ <- factor(msPHATE_clusters)
bcn$msphate_subclusters_lvl4_scaled <- msPHATE_clusters
```

```{r}
write.table(bcn@misc$scale.data_unscaled, file = paste(wdir, 'counts.csv', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(colnames(bcn@misc$scale.data_unscaled), file = paste(wdir, 'cell_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(rownames(bcn@misc$scale.data_unscaled), file = paste(wdir, 'gene_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
# Run msPHATE
```

```{python}
### RESTART ###
msphate(4)
# Use msphate.py
```

```{r}
msPHATE_embedding <- read.csv(paste(wdir, 'msPHATE_embedding.csv', sep = ''), header=FALSE, row.names = 1, col.names = c('', 'msPHATE_1', 'msPHATE_2'))
msPHATE_embedding <- as.matrix(msPHATE_embedding)
bcn[['msphate_unscaled']] <- CreateDimReducObject(embeddings = msPHATE_embedding, key = 'msPHATE_')
```

```{r}
msPHATE_clusters <- as.matrix(read.csv(paste(wdir, 'msPHATE_clusters.csv', sep = ''), header=FALSE))
msPHATE_clusters <- factor(msPHATE_clusters)
bcn$msphate_subclusters_lvl4_unscaled <- msPHATE_clusters
```

```{r}
bcn <- RunPCA(object = bcn, reduction.name = 'pca_scaled')
bcn <- FindNeighbors(object = bcn, reduction = 'pca_scaled')
bcn <- FindClusters(object = bcn, resolution = 1.2)
bcn <- RunUMAP(object = bcn, features = VariableFeatures(bcn))
bcn <- RunTSNE(object = bcn, features = VariableFeatures(bcn), reduction = 'pca_scaled')
```


### Using corresponding msPHATE clusters

```{r}
write.table(GetAssayData(bcn_ms, 'scale.data'), file = paste(wdir, 'counts.csv', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(colnames(GetAssayData(bcn_ms, 'scale.data')), file = paste(wdir, 'cell_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(rownames(GetAssayData(bcn_ms, 'scale.data')), file = paste(wdir, 'gene_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
# Run msPHATE
```

```{python}
### RESTART ###
msphate(4)
# Use msphate.py
```

```{r}
msPHATE_embedding <- read.csv(paste(wdir, 'msPHATE_embedding.csv', sep = ''), header=FALSE, row.names = 1, col.names = c('', 'msPHATE_1', 'msPHATE_2'))
msPHATE_embedding <- as.matrix(msPHATE_embedding)
bcn_ms[['msphate_scaled']] <- CreateDimReducObject(embeddings = msPHATE_embedding, key = 'msPHATE_')
```

```{r}
msPHATE_clusters <- as.matrix(read.csv(paste(wdir, 'msPHATE_clusters.csv', sep = ''), header=FALSE))
msPHATE_clusters_ <- factor(msPHATE_clusters)
bcn_ms$msphate_subclusters_lvl4_scaled <- msPHATE_clusters
```

```{r}
write.table(bcn_ms@misc$scale.data_unscaled, file = paste(wdir, 'counts.csv', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(colnames(bcn_ms@misc$scale.data_unscaled), file = paste(wdir, 'cell_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
write.table(rownames(bcn_ms@misc$scale.data_unscaled), file = paste(wdir, 'gene_names.txt', sep = ''), sep = ',', row.names = FALSE, col.names = FALSE)
# Run msPHATE
```

```{python}
### RESTART ###
msphate(4)
# Use msphate.py
```

```{r}
msPHATE_embedding <- read.csv(paste(wdir, 'msPHATE_embedding.csv', sep = ''), header=FALSE, row.names = 1, col.names = c('', 'msPHATE_1', 'msPHATE_2'))
msPHATE_embedding <- as.matrix(msPHATE_embedding)
bcn_ms[['msphate_unscaled']] <- CreateDimReducObject(embeddings = msPHATE_embedding, key = 'msPHATE_')
```

```{r}
msPHATE_clusters <- as.matrix(read.csv(paste(wdir, 'msPHATE_clusters.csv', sep = ''), header=FALSE))
msPHATE_clusters <- factor(msPHATE_clusters)
bcn_ms$msphate_subclusters_lvl4_unscaled <- msPHATE_clusters
```

```{r}
bcn_ms <- RunPCA(object = bcn_ms, reduction.name = 'pca_scaled')
bcn_ms <- FindNeighbors(object = bcn_ms, reduction = 'pca_scaled')
bcn_ms <- FindClusters(object = bcn_ms, resolution = 1.2)
bcn_ms <- RunUMAP(object = bcn_ms, features = VariableFeatures(bcn_ms))
bcn_ms <- RunTSNE(object = bcn_ms, features = VariableFeatures(bcn_ms), reduction = 'pca_scaled')
```

```{r}
saveRDS(bcn, file = paste(wdir, 'bcn_final.RDS', sep = ''))
saveRDS(bcn_ms, file = paste(wdir, 'bcn_ms_final.RDS', sep = ''))
```

```{r}
DimPlot(bcn, reduction = 'msphate_scaled', group.by = 'jansky_idents', pt.size = 1) + ggtitle('msphate_scaled by jansky_idents')
DimPlot(bcn, reduction = 'msphate_scaled', group.by = 'msphate_subclusters_lvl4_scaled', pt.size = 1) + ggtitle('msphate_scaled at lvl4')
DimPlot(bcn, reduction = 'msphate_unscaled', group.by = 'jansky_idents', pt.size = 1) + ggtitle('msphate_unscaled by jansky_idents')
DimPlot(bcn, reduction = 'msphate_unscaled', group.by = 'msphate_subclusters_lvl4_unscaled', pt.size = 1) + ggtitle('msphate_unscaled at lvl4')

## Minor - fix umap and tsne naming in bcn
## Add label = TRUE? For equal sizing
DimPlot(bcn, reduction = 'umap', group.by = 'jansky_idents', pt.size = 1) + ggtitle('umap_scaled by jansky_idents')
DimPlot(bcn, reduction = 'umap', group.by = 'seurat_clusters', pt.size = 1) + ggtitle('umap_scaled by seurat_clusters')
DimPlot(bcn, reduction = 'umap_unscaled', group.by = 'jansky_idents', pt.size = 1) + ggtitle('umap_unscaled by jansky_idents')
DimPlot(bcn, reduction = 'umap_unscaled', group.by = 'seurat_clusters', pt.size = 1) + ggtitle('umap_unscaled by seurat_clusters')

DimPlot(bcn, reduction = 'tsne', group.by = 'jansky_idents', pt.size = 1) + ggtitle('tsne_scaled by jansky_idents')
DimPlot(bcn, reduction = 'tsne', group.by = 'seurat_clusters', pt.size = 1) + ggtitle('tsne_scaled by seurat_clusters')
DimPlot(bcn, reduction = 'tsne_unscaled', group.by = 'jansky_idents', pt.size = 1) + ggtitle('tsne_unscaled by jansky_idents')
DimPlot(bcn, reduction = 'tsne_unscaled', group.by = 'seurat_clusters', pt.size = 1) + ggtitle('tsne_unscaled by seurat_clusters')
```

```{r}
DimPlot(bcn_ms, reduction = 'msphate_scaled', group.by = 'jansky_idents', pt.size = 1) + ggtitle('msphate_scaled by jansky_idents')
DimPlot(bcn_ms, reduction = 'msphate_scaled', group.by = 'msphate_subclusters_lvl4_scaled', pt.size = 1) + ggtitle('msphate_scaled at lvl4')
DimPlot(bcn_ms, reduction = 'msphate_unscaled', group.by = 'jansky_idents', pt.size = 1) + ggtitle('msphate_unscaled by jansky_idents')
DimPlot(bcn_ms, reduction = 'msphate_unscaled', group.by = 'msphate_subclusters_lvl4_unscaled', pt.size = 1) + ggtitle('msphate_unscaled at lvl4')

## Minor - fix umap and tsne naming in bcn_ms
DimPlot(bcn_ms, reduction = 'umap', group.by = 'jansky_idents', pt.size = 1) + ggtitle('umap_scaled by jansky_idents')
DimPlot(bcn_ms, reduction = 'umap', group.by = 'seurat_clusters', pt.size = 1) + ggtitle('umap_scaled by seurat_clusters')
DimPlot(bcn_ms, reduction = 'umap_unscaled', group.by = 'jansky_idents', pt.size = 1) + ggtitle('umap_unscaled by jansky_idents')
DimPlot(bcn_ms, reduction = 'umap_unscaled', group.by = 'seurat_clusters', pt.size = 1) + ggtitle('umap_unscaled by seurat_clusters')

DimPlot(bcn_ms, reduction = 'tsne', group.by = 'jansky_idents', pt.size = 1) + ggtitle('tsne_scaled by jansky_idents')
DimPlot(bcn_ms, reduction = 'tsne', group.by = 'seurat_clusters', pt.size = 1) + ggtitle('tsne_scaled by seurat_clusters')
DimPlot(bcn_ms, reduction = 'tsne_unscaled', group.by = 'jansky_idents', pt.size = 1) + ggtitle('tsne_unscaled by jansky_idents')
DimPlot(bcn_ms, reduction = 'tsne_unscaled', group.by = 'seurat_clusters', pt.size = 1) + ggtitle('tsne_unscaled by seurat_clusters')
```
